---
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE)

## The following document outputs a visual for the altadena_housing_before_and_after.html 

#### Setting Up ####

library(usethis)
library(dplyr)
library(tidyverse)
library(htmltools) 
library(htmlwidgets)
library(gtExtras)
library(gt)
library(sf)
library(scales)
library(leaflet)
library(leaflet.extras)
library(rmapshaper)

options(scipen=999)

source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("altadena_recovery_rebuild")

## COLORS ##
orange <- "#F25922"
peridot <- "#CEEA01"
lavender <- "#CEC4E2"
meteorite <- "#3A207D"
ccblue <- "#0860BC"

black <- "#000000"
alabaster<-"#FBFBFB"
gainsboro <- "#DEDEDE"
```

```{r}
# Eaton Fire damaged properties
# residential parcel shapes
parcels <- st_read(con, query="SELECT * FROM data.rel_assessor_altadena_parcels_sept2025", geom="geom") %>% 
  st_transform(4326)

res_data <- st_read(con, query="SELECT * FROM data.rel_assessor_residential_sept2025")
damage <- st_read(con, query="SELECT * FROM data.rel_assessor_damage_level_sept2025")

# join parcel shapes to residential data
parcels <- parcels %>% left_join(res_data, by="ain_sept") %>% left_join(damage, by="ain_sept")

# make one layer for all residential - use this to visualize all properties
parcels_res_damaged <- parcels %>% 
  filter((residential==TRUE & total_units>0 & damage_category=="Significant Damage"))

# east and west altadena boundaries --
west <- st_read(con, query="SELECT * FROM data.west_altadena_3310", geom="geom") %>% 
  st_transform(4326) %>% ms_simplify()

east <- st_read(con, query="SELECT * FROM data.east_altadena_3310", geom="geom") %>% 
  st_transform(4326) %>%  ms_simplify()

df_lead <- st_read(con, query="Select * from data.lacdph_lead_results_grid_2025", geom="geom") %>% 
  st_transform(4326) %>%
  ms_simplify() %>%
  mutate(hi_lead_flag = case_when(is.na(lead_geometric_mean) ~ "Not Tested",
                               lead_geometric_mean >= 80 ~ "Above Threshold",
                               lead_geometric_mean <80 ~ "Below Threshold",
                               TRUE ~ NA)) 

#NOTE: taking time to transform even after filtering
df_zoning_lac <- st_read(con, query = "Select * FROM data.lac_unincorporated_zoning_3310 WHERE plng_area = 'West San Gabriel Valley'", geom="geom") %>% st_transform(4326) %>% ms_simplify()

df_local_fhsz <- st_read(con, query="Select * from data.eaton_local_fhsz_4326 WHERE fhsz_descr != 'NonWildland'", geom="geom") 

df_state_fhsz <- st_read(con, query="Select * from data.eaton_state_fhsz_4326", geom="geom")

eaton_fire <- st_read(con, query="SELECT * FROM data.eaton_fire_prmtr_3310", geom="geom") %>% st_transform(4326) %>%  ms_simplify()

#Pulling in data for POP UP 
sept_zoning <- st_read(con, query="SELECT * FROM data.rel_assessor_zoning_sept2025")
sept_fhsz <- st_read(con, query="SELECT * FROM data.rel_assessor_fhsz_sept2025") %>% select(-area_name)
sept_lead <- st_read(con, query="SELECT * FROM data.rel_assessor_lead_sept2025")
sept_sales <- dbGetQuery(con, statement = "SELECT ain_sept, sold_after_eaton, first_owner_name, first_owner_name_overflow, second_owner_name FROM data.rel_assessor_sales_sept2025;") %>%
  mutate(owner_names = str_squish(paste(first_owner_name, first_owner_name_overflow, second_owner_name))) %>%
  select(ain_sept, sold_after_eaton, owner_names)
sept_rebuild <- dbGetQuery(con, statement="SELECT ain as ain_jan, ain_sept, address_sept, rebuild_status
FROM data.rel_parcel_rebuild_status_2025_10
WHERE damage_category='Significant Damage';") %>%
  mutate(rebuild_status = str_to_title(rebuild_status))
sept_contact <- dbGetQuery(con, "SELECT ain as ain_sept, mail_house_no, m_fraction, m_direction, m_street_name, m_unit, m_city_state, m_zip FROM data.assessor_data_universe_sept2025;") %>%
  # create clean address for contact
  mutate(
    street_part = trimws(paste(mail_house_no, m_fraction, m_direction, m_street_name)),
    # Add unit if present
    unit_part = ifelse(!is.na(m_unit) & trimws(m_unit) != "", paste0(", ", trimws(m_unit)), ""),
    city_zip_part = paste0(m_city_state, " ", substr(m_zip, 1, 5)),
    # Combine everything
    contact_address = str_squish(paste0(street_part, unit_part, ", ", city_zip_part))) %>%
  select(ain_sept, contact_address) %>% 
  filter(ain_sept %in% parcels_res_damaged$ain_sept)

all_df <- parcels_res_damaged %>% 
  left_join(sept_zoning, by = "ain_sept") %>% 
  left_join(sept_fhsz, by = "ain_sept") %>% 
  left_join(sept_lead, by="ain_sept") %>%
  left_join(sept_sales, by="ain_sept") %>%
  left_join(sept_contact, by = "ain_sept") %>%
  left_join(sept_rebuild, by="ain_sept")


### factors and paletes for other polygons ----
# high lead grid colors
df_lead$hi_lead_flag <- factor(
      df_lead$hi_lead_flag,
      levels = c("Not Tested","Below Threshold","Above Threshold"))

# Create a color palette based on the factor
lead_pal <- colorFactor(
  palette = c("gray","#ffcccc","red"), 
  domain = df_lead$hi_lead_flag)

# fire zone colors
df_local_fhsz$fhsz_descr <- factor(
      df_local_fhsz$fhsz_descr,
      levels = c("Moderate","High","Very High"))

# Create a color palette based on the factor
colorpal_local_fhsz  <- colorFactor(palette = c("#ffffb2","#fd8d3c","#f03b20"), domain = df_local_fhsz$fhsz_descr)


# zone colors
# reference https://planning.lacounty.gov/wp-content/uploads/2022/04/map_z_02_Altadena.pdf
# table(df_zoning_lac$z_desc,useNA='always')
zoning_codes <- as.data.frame(table(df_zoning_lac$z_category,useNA='always'))
df_zoning_lac <- df_zoning_lac %>%
  mutate(zone_label=
           case_when(
             str_detect(z_category, "^C") ~ "Commercial",
             str_detect(z_category, "^A") ~ "Agricultural",
             str_detect(z_category, "^R-1") ~ "Single-Family Residence",
             str_detect(z_category, "^R-2") ~ "Two-Family Residence",
             str_detect(z_category, "^R-3") ~ "Multiple Residence",
             str_detect(z_category, "^R-4") ~ "Multiple Residence",
             str_detect(z_category, "^R-5") ~ "Multiple Residence",
             str_detect(z_category, "^R") ~ "Other Residential",
             str_detect(z_category, "^MXD") ~ "Mixed Use Development",
             str_detect(z_category, "^M-") ~ "Manufacturing",
             str_detect(z_category, "^O-") ~ "Open Space/Water Shed",
             str_detect(z_category, "^W-") ~ "Open Space/Water Shed",
             TRUE ~ "Other"))
# check
check <- df_zoning_lac %>% st_drop_geometry() %>% group_by(zone_label, z_category) %>% summarise(count=n())
             
df_zoning_lac$zone_label<- factor(
      df_zoning_lac$zone_label,
      levels = c("Single-Family Residence","Two-Family Residence","Multiple Residence","Other Residential","Mixed Use Development","Commercial","Manufacturing","Agricultural","Open Space/Water Shed","Other"))

# parcels sold after eaton fire
sept_sales_sf <- all_df %>%
  filter(sold_after_eaton==TRUE) %>%
  select(ain_sept, sold_after_eaton, geom) 

# rebuild status
sept_rebuild_sf <- all_df %>%
  select(ain_sept, rebuild_status) %>%
  mutate(rebuild_status_factor = factor(
    rebuild_status, 
    levels=c("Debris Removal Incomplete", "Awaiting Permit Application", "Awaiting Construction Start", "Construction In Progress", "Rebuild Complete")))

# Create a color palette based on the factor
zoning_pal <- colorFactor(
  palette = c("#feffe2","#f2d84b","#d09a36","#dfd3b6","#b3605d","#E34234","#b7c4db","#c5fb5d","#8ffa4f","#bfbfbf"), 
  domain = df_zoning_lac$zone_label)

rebuild_pal <- colorFactor(
  palette = c("#E3D5FF", "#C3A0F4", "#8B5FBF", "#4C2393", "#2E0C6A"),
  domain= sept_rebuild_sf$rebuild_status_factor
)

```

```{r}
# final columns 
all_df <- all_df %>%
  select(ain_sept,
         address_sept,
         res_type,
         owner_renter,
         owner_names,
         contact_address,
         total_square_feet,
         total_units,
         z_desc,
         combined_fhsz,
         hi_lead_label,
         rebuild_status,
         sold_after_eaton)

popup_parcel <- paste0("<div class='parcel-popups'> <div class='leaflet-data-parcel' data-ain='", all_df$ain_sept,
               "' data-parcel-address='", all_df$address_sept,
               "' data-res-type='", all_df$res_type,
               "' data-owner-type='", all_df$owner_renter,
               "' data-owner-names='", all_df$owner_names,
               "' data-owner-contact='", all_df$contact_address,
               "' data-total-sqft='", comma(all_df$total_square_feet),
               "' data-total-units='", all_df$total_units,
               "' data-lac-zoning='", all_df$z_desc,
               "' data-fire-zone='", all_df$combined_fhsz, 
               "' data-lead-level='", all_df$hi_lead_label, 
               "' data-rebuild-status='", all_df$rebuild_status, 
               "' data-sold-status='", all_df$sold_after_eaton, 
               "'></div></div>")

## Create custom popup
# We are using <div> "class" names (e.g. popup-body) to apply CSS styling (end of script, before JS)
# We are using <span> "class" names (e.g. res-type) to style with CSS AND insert the data stored in data_popup
# Note: the <span> "class" names are using the SAME naming conventions as in the data_popup (only difference is no "data-" prefix) - this makes things clear and convenient for JS coding (end of script)

custom_sidebar <- paste0("<div class='leaflet-popup-scrolled parcel-sidebar'>",
                              "<div class='popup-instruction'>Click on an Altadena parcel to learn more.</div>",
                              "<div class ='geo-header'>", 
                              "<div class='popup-body'>",
                              "AIN: <span class='ain'></span>",
                              "<br>",
                              "Parcel Address: <span class='parcel-address'></span>",
                              "<br>",
                              "Residential Type: <span class='res-type'></span>",
                              "<br>",
                              "Ownership Type: <span class='owner-type'></span>",
                              "<br>",
                              "Owner Name(s): <span class='owner-names'></span>",
                              "<br>",
                              "Owner Contact: <span class='owner-contact'></span>",
                              "<br>",
                              "Total Square Footage: <span class='total-sqft'></span>",
                              "<br>",
                              "Total Units: <span class='total-units'></span>",
                              "<br>",
                              "L.A. County Zoning: <span class='lac-zoning'></span>",
                              "<br>",
                              "Fire Hazard Zone: <span class='fire-zone'></span>",
                              "<br>",
                              "Lead Level: <span class='lead-level'></span>",
                              "<br>",
                              "Rebuild Status: <span class='rebuild-status'></span>",
                              "<br>",
                              "Sold After Eaton: <span class='sold-status'></span>",
                              "<br>","</div>",
                              "</div>",
                              "</div>")


```


```{r}

leaflet(
  width = "100%",
  height = "600px",
  options = leafletOptions(zoomSnap = 0.1, zoomDelta = 0.1,
                           attributionControl=FALSE)) %>%
  setView(-118.137, 34.194, zoom = 13) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  
  # add custom sidebar
  addControl(html=custom_sidebar, position = "topleft") %>%
     
  #Damaged/Destroyed Properties
  addPolygons(
    data=all_df,
    fillColor="orange",
    fillOpacity = 1,
    color="white",
    weight=.4,
    opacity=1,
    popup = popup_parcel,
    group="Significantly Damaged Parcels") %>%
  #Eaton Fire Perimeter
  addPolygons(
    data=eaton_fire,
    fillColor="white",
    fillOpacity=0,
    color="black",
    weight=2.5,
    opacity=1,
    options=pathOptions(clickable=FALSE),
    group="Eaton Fire Perimeter") %>%
  #East Perimeter
  addPolygons(
    data=east,
    fillColor="white",
    fillOpacity=0,
    color=meteorite,
    weight=2.5,
    opacity=1,
    options=pathOptions(clickable=FALSE),
    group="East Altadena") %>%
  #West Perimeter
  addPolygons(
    data=west,
    fillColor="white",
    fillOpacity=0,
    color=meteorite,
    weight=2.5,
    opacity=1,
    options=pathOptions(clickable=FALSE),
    group="West Altadena") %>%
  #Lead Data 
  addPolygons(
    data=df_lead,
    fillColor = ~lead_pal(hi_lead_flag),
    fillOpacity=0.4,
    color="white",
    weight=.5,
    opacity=1,
    popup=~paste0("Average Lead Level Measured in mg/kg: ", 
                  df_lead$lead_geometric_mean, "</br>",
                  "Higher than 80 mg/kg Safety Threshold: ", df_lead$hi_lead_flag),
    group="Lead Levels in Soil") %>%
  addLegend(
    position = "bottomright",
    pal = lead_pal,
    values =  df_lead$hi_lead_flag,
    group="Lead Levels in Soil",
    title = "Lead Levels",
    opacity=0.4) %>%
  #Fire Hazard Zones 
  addPolygons(
    data=df_local_fhsz,
    fillColor=~colorpal_local_fhsz(fhsz_descr),
    fillOpacity=0.5,
    color="white",
    weight=1,
    popup=~fhsz_descr,
    group="Local Fire Hazard Zones") %>%
  addLegend(
    position = "bottomright",
    pal = colorpal_local_fhsz,
    values =  df_local_fhsz$fhsz_descr,
    group="Local Fire Hazard Zones",
    title = "Fire Hazard Level",
    opacity=0.5) %>%
  addPolygons(
    data=df_state_fhsz,
    fillColor="red",
    fillOpacity=0.5,
    color="white",
    weight=1,
    popup=~fhsz_descr,
    group="State Fire Hazard Zones") %>%
  #Zoning 
  addPolygons(
    data=df_zoning_lac,
    fillColor=~zoning_pal(zone_label),
    fillOpacity=0.5,
    color="white",
    weight=1,
    opacity=1,
    popup=~paste0("Zoning Description: ", df_zoning_lac$z_desc),
    group="Zoning") %>%
  addLegend(
    position = "bottomright",
    pal = zoning_pal,
    values =  df_zoning_lac$zone_label,
    group="Zoning",
    title = "Allowable Uses",
    opacity=0.5) %>% 
  # Sold Status
  addPolygons(
    data=sept_sales_sf,
    fillColor="black",
    fillOpacity=1,
    color="white",
    weight=0.2,
    opacity=1,
    popup=popup_parcel,
    group="Sold After Eaton Fire") %>%
  # Rebuild Status
  addPolygons(
    data=sept_rebuild_sf,
    fillColor=~rebuild_pal(rebuild_status_factor),
    fillOpacity=1,
    color="white",
    weight=0.4,
    opacity=1,
    popup=popup_parcel,
    group="Rebuild Status") %>%
  addLegend(
    position = "bottomright",
    pal = rebuild_pal,
    values =  sept_rebuild_sf$rebuild_status_factor,
    group="Rebuild Status",
    title = "Rebuild Status",
    opacity=1) %>% 
  
  #Controling Layers
  addLayersControl(
    overlayGroups = c("Significantly Damaged Parcels",
                      "Lead Levels in Soil","Local Fire Hazard Zones",
                      "State Fire Hazard Zones","Zoning", "Sold After Eaton Fire",
                      "Rebuild Status", "West Altadena","East Altadena",
                      "Eaton Fire Perimeter"),
    options = layersControlOptions(collapsed = FALSE, autoZIndex = TRUE))  %>%
  hideGroup(c("Eaton Fire Perimeter","Lead Levels in Soil","Local Fire Hazard Zones",
              "State Fire Hazard Zones","Zoning", "Sold After Eaton Fire", "Rebuild Status")) %>% addFullscreenControl()

```

<style type="text/css">

/* minimum required css for leaflet maps with custom sidebars */ 

.info {
    padding: 6px 8px;
    font-size: 12px;
    background: white;
    background: rgba(255,255,255,0.8);
    border-radius: 5px;
    color: #000000;
}

.leaflet-touch .leaflet-control-layers {
  border-left:1px solid #737373;
  border-right:1px solid #737373;
  border-top:1px solid #737373;
  border-bottom:1px solid #737373;
}
/* Target all leaflet controls on the left side, regardless of top/bottom position */
/* when keeping zoom buttons, remember to exclude from this styling */
.leaflet-left .leaflet-control:not(.leaflet-control-zoom) {
  background-color: white;
  opacity: 1.0;
  overflow: auto;
  padding: 2%;
  height: 100%;
  margin: 2px;
  width: 200px;
}


/* Ensure consistent styling for popup content */
.leaflet-popup-scrolled {
  overflow: visible !important;
  max-height: none !important;
  border-bottom: none !important;
  border-top: none !important;
}

/* Ensure the outer container has scrolling enabled */
.info.leaflet-control {
  overflow: auto !important;
  max-height: 596px !important;
}

.leaflet-right .leaflet-control {
  background-color: white;
  opacity: 1.0;
  padding: 2%;
  margin: 2px;
  width: 200px;
}

.leaflet-control-layers .leaflet-control .leaflet-control-layers-expanded {
  background-color: white;
  opacity: 1.0;
}

.hidden {
  display: none;
}

/* css styling for headers */
.popup-body {
  color: #000000;
  font-size: 14px;
  font-weight: bold;
}

/* css styling that can change depending on the map data and custom styling needs */

.popup-instruction {
  font-style: italic;
  font-weight: 600; 
  line-height: 16px;
  color: #000000;
  border-bottom: 1px solid #000000;
  padding-bottom: 16px;
  margin-bottom: 16px;
}

.ain,
.parcel-address,
.res-type,
.owner-type,
.owner-names,
.owner-contact,
.total-sqft,
.total-units,
.lac-zoning,
.fire-zone,
.lead-level,
.rebuild-status,
.sold-status {
  font-weight: normal; 
}


</style>

```{js}
// This function is used to format WRCA Congressional District data in the chloropleth map "popup"/sidebar feature
// Standard set up for using JS with a leaflet map
var mapsPlaceholder = [];
L.Map.addInitHook(function () {
  mapsPlaceholder.push(this);
  mapsPlaceholder.forEach(map => {

      // to pre-populate map pop-up (when map loads/before user clicks anywhere)
      map.on('load', (e) => {
        setTimeout(() => {
          document.querySelector('.ain').innerHTML = "5833017027"
          document.querySelector('.parcel-address').innerHTML = "140 E LAS FLORES DR ALTADENA CA"
          document.querySelector('.res-type').innerHTML = "Single-family"
          document.querySelector('.owner-type').innerHTML = "Renter occupied"
          document.querySelector('.owner-names').innerHTML = "PEREZ,FERNANDO AND VIOLETA TRS PEREZ FAMILY TRUST"
          document.querySelector('.owner-contact').innerHTML = "140 E LAS FLORES DR, ALTADENA CA 91001"
          document.querySelector('.total-sqft').innerHTML = "1380"
          document.querySelector('.total-units').innerHTML = "1"
          document.querySelector('.lac-zoning').innerHTML = "Single-Family Residence"
          document.querySelector('.fire-zone').innerHTML = "None"
          document.querySelector('.lead-level').innerHTML = "Not High Lead Grid"
          document.querySelector('.rebuild-status').innerHTML = "Construction In Progress"
          document.querySelector('.sold-status').innerHTML = "FALSE"
          
          
        }, 500) //assumes the map will load within 500ms
      }) // end map load check

      // hide default leaflet popup
      map.on('popupopen', (e) => {
        let popupsParcel = document.querySelectorAll('.leaflet-data-parcel');

        document.querySelectorAll('.parcel-popups').forEach(parcelPopup => {
          parcelPopup.parentElement.parentElement.parentElement.classList.add('hidden')
        })

        // populate custom pop-up when user clicks map
        let popupParcel = popupsParcel[popupsParcel.length-1]
        document.querySelector('.ain').innerHTML = popupParcel.getAttribute('data-ain')
        document.querySelector('.parcel-address').innerHTML = popupParcel.getAttribute('data-parcel-address')
        document.querySelector('.res-type').innerHTML = popupParcel.getAttribute('data-res-type')
        document.querySelector('.owner-type').innerHTML = popupParcel.getAttribute('data-owner-type')
        document.querySelector('.owner-names').innerHTML = popupParcel.getAttribute('data-owner-names')
        document.querySelector('.owner-contact').innerHTML = popupParcel.getAttribute('data-owner-contact')
        document.querySelector('.total-sqft').innerHTML = popupParcel.getAttribute('data-total-sqft')
        document.querySelector('.total-units').innerHTML = popupParcel.getAttribute('data-total-units')
        document.querySelector('.lac-zoning').innerHTML = popupParcel.getAttribute('data-lac-zoning')
        document.querySelector('.fire-zone').innerHTML = popupParcel.getAttribute('data-fire-zone')
        document.querySelector('.lead-level').innerHTML = popupParcel.getAttribute('data-lead-level')
        document.querySelector('.rebuild-status').innerHTML = popupParcel.getAttribute('data-rebuild-status')
        document.querySelector('.sold-status').innerHTML = popupParcel.getAttribute('data-sold-status')
      }) // end popupopen check
  })
});

```


```{r}
dbDisconnect(con)
```