---
title: ""
output: rmarkdown::html_vignette
# css: "W:\\RDA Team\\R\\cc_brandguide.css"

---

```{r, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE
)

## The following document outputs a visual of the housing stock before in a table format for the altadena_housing_before_and_after.html 

#### Setting Up ####

library(usethis)
library(dplyr)
library(tidyverse)
library(htmltools) 
library(htmlwidgets)
library(gtExtras)
library(gt)
library(sf)
library(scales)

options(scipen=999)

source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("altadena_recovery_rebuild")

## COLORS ##
orange <- "#F25922"
peridot <- "#CEEA01"
lavender <- "#CEC4E2"
meteorite <- "#3A207D"
ccblue <- "#0860BC"

black <- "#000000"
alabaster<-"#FBFBFB"
gainsboro <- "#DEDEDE"

```

In January 2025, across Altadena, multifamily homes had the highest average square footage at 2,911 square feet. Overall, East Altadena homes had a higher average square footage than those in West Altadena with the exception of condominiums. Condominiums in West Altadena on average had 300 more square footage than the ones in East Altadena. In West Altadena, the average single-family home was 1,571 square feet.

```{r}
#### VISUAL ####
df_ <- st_read(con, query="Select res_type, area, sqftg_avg, sqftg_sum from data.analysis_sqftg_jan2025") 

df <- df_ %>%
  filter(res_type != "Boarding house") %>% 
  mutate(" "=paste0(comma(round(sqftg_avg,0))," sqft"),
         "Average Square Footage"=sqftg_avg,
         "Residential Type" = res_type,
         "Area" = str_to_title(area)) %>%
  mutate("Area" = factor(Area, levels = c("Altadena", "West", "East"))) %>%
  arrange(desc(sqftg_sum)) 


table<- df %>%
  select("Area","Residential Type", "Average Square Footage", " ")%>%
  gt(groupname_col = "Area", rowname_col = "Residential Type") %>%
  tab_footnote(footnote=html("Catalyst California's calculations based on L.A. County Assessor Data, January 2025."),placement="left") 

 table%>%
  gt_plt_bar_pct(column= 'Average Square Footage', scaled=FALSE, fill=meteorite, background="transparent")%>%
  # if needed modify column widths here 
  cols_width(1~px(210),2~px(100),
             3~px(150))%>%
  opt_align_table_header("left")%>%
  opt_table_font(font=c("Inter", default_fonts())) %>% # opt_table_font(font=c(google_font(name="Inter"),locations=NULL,default_fonts()))%>%
  tab_options(table.background.color=alabaster,
    column_labels.background.color = "transparent",
    table.font.size = px(12),
    column_labels.font.size = px(12),
    heading.align = "left",
    table.font.names = "Inter",
    heading.title.font.size = px(20),
    heading.border.bottom.color = "transparent",
    column_labels.border.bottom.width = px(3),
    column_labels.border.bottom.color = gainsboro,
    table.border.top.color = "transparent",
    table.border.bottom.color = "transparent",
    table.border.bottom.width = px(3),
    data_row.padding = px(3),
    footnotes.font.size = px(10),
    footnotes.border.bottom.color = "transparent",
    table.align='left')%>%
  tab_style(style=
            cell_borders(sides=c("top","bottom"),color=gainsboro,style="solid",weight=px(1)),locations=cells_body())%>%
   tab_style(style=
            cell_text(font=c("Inter", default_fonts()),weight=900),locations=cells_title()) # cell_text(font=c(google_font(name="Inter"),default_fonts()),weight=900),locations=cells_title())

```

```{r, include=FALSE}
dbDisconnect(con) 
```