---
title: ""
output: rmarkdown::html_vignette
# css: "W:\\RDA Team\\R\\cc_brandguide.css"

---

```{r, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE
)

## The following document outputs a visual of the housing stock before in a table format for the altadena_housing_before_and_after.html 

#### Setting Up ####

library(usethis)
library(dplyr)
library(tidyverse)
library(htmltools) 
library(htmlwidgets)
library(gtExtras)
library(gt)
library(sf)
library(scales)

options(scipen=999)

source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("altadena_recovery_rebuild")

## COLORS ##
orange <- "#F25922"
peridot <- "#CEEA01"
lavender <- "#CEC4E2"
meteorite <- "#3A207D"
ccblue <- "#0860BC"

black <- "#000000"
alabaster<-"#FBFBFB"
gainsboro <- "#DEDEDE"

```

To understand the types of rental units lost, we then focused on all properties with rental units recorded. We examined what percentage of rental units were lost by residential type. Across Altadena and residential housing types, about one in three rental units were lost in significantly damaged or destroyed properties. West Altadena rental units were hit the hardest. Nearly half of rental units in multifamily properties were significant damaged or destroyed in the fire. And over 40% of rental units in single-family homes in West Altadena were significantly damaged or destroyed. In order to return Altadena at least to the level of rental options existing before the fire, significant investments and strategic solutions to rebuilding rental properties is required, especially as renters are more vulnerable to displacement after disaster.

```{r}
#### VISUAL ####
df_ <- st_read(con, query="Select res_type, damage_category, area_name, rent_units_count, rent_units_total, rent_units_prc from data.analysis_units_damage") 

df <- df_ %>%
  filter(res_type != "Boarding house",
         damage_category == "Significant Damage") %>% 
  mutate("% (Count)"=paste0(round(rent_units_prc,0),"% (", comma(rent_units_count)," out of ", 
                            comma(rent_units_total), ")"),
         
         "Rental Units Lost in Housing Type"=rent_units_prc,
         "Residential Type" = res_type,
         "Area" = str_to_title(area_name)) %>%
  mutate("Area" = factor(Area, levels = c("Altadena", "West", "East"))) %>%
  arrange(desc(rent_units_count)) 


table<- df %>%
  select("Area","Residential Type", "Rental Units Lost in Housing Type", "% (Count)")%>%
  gt(groupname_col = "Area", rowname_col = "Residential Type") %>%
  tab_header(title= "Significantly Damaged Rental Units by Housing Type")%>%
  tab_footnote(footnote=html("Catalyst California's calculations based on L.A. County Assessor Data, January 2025 and CAL FIRE Damage Data, September 2025."),placement="left") 

 table%>%
  gt_plt_bar_pct(column= 'Rental Units Lost in Housing Type', scaled=TRUE, fill=meteorite, background=gainsboro)%>%
  # if needed modify column widths here 
  # cols_width(1~px(210),2~px(100),
  #            3~px(125))%>%
  opt_align_table_header("left")%>%
  opt_table_font(font=c("Inter", default_fonts())) %>% # opt_table_font(font=c(google_font(name="Inter"),locations=NULL,default_fonts()))%>%
  tab_options(table.background.color=alabaster,
    column_labels.background.color = "transparent",
    table.font.size = px(12),
    column_labels.font.size = px(12),
    heading.align = "left",
    table.font.names = "Inter",
    heading.title.font.size = px(20),
    heading.border.bottom.color = "transparent",
    column_labels.border.bottom.width = px(3),
    column_labels.border.bottom.color = gainsboro,
    table.border.top.color = "transparent",
    table.border.bottom.color = "transparent",
    table.border.bottom.width = px(3),
    data_row.padding = px(3),
    footnotes.font.size = px(10),
    footnotes.border.bottom.color = "transparent",
    table.align='left')%>%
  tab_style(style=
            cell_borders(sides=c("top","bottom"),color=gainsboro,style="solid",weight=px(1)),locations=cells_body())%>%
   tab_style(style=
            cell_text(font=c("Inter", default_fonts()),weight=900),locations=cells_title()) # cell_text(font=c(google_font(name="Inter"),default_fonts()),weight=900),locations=cells_title())


```


```{r, include=FALSE}
dbDisconnect(con) 
```