---
title: ""
output: rmarkdown::html_vignette
# css: "W:\\RDA Team\\R\\cc_brandguide.css"

---

```{r, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE
)

## The following document outputs a visual of the housing stock before in a table format for the altadena_housing_before_and_after.html 

#### Setting Up ####

library(usethis)
library(dplyr)
library(tidyverse)
library(htmltools) 
library(htmlwidgets)
library(gtExtras)
library(gt)
library(sf)
library(scales)

options(scipen=999)

source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("altadena_recovery_rebuild")

## COLORS ##
orange <- "#F25922"
peridot <- "#CEEA01"
lavender <- "#CEC4E2"
meteorite <- "#3A207D"
ccblue <- "#0860BC"

black <- "#000000"
alabaster<-"#FBFBFB"
gainsboro <- "#DEDEDE"

```

We also examined the share of each ownership type that has sold since the Eaton Fire. Understanding which properties are selling at faster rates can help signal trends toward displacement and disaster capitalism, or corporate investors exploiting survivors during their recovery.^[https://www.saje.net/wp-content/uploads/2025/10/ConfrontingDisaster.10.10.2025.pdf] As of September 2025, most properties that have sold since the fire were owner-occupied before fire, but these properties are not selling at a faster rate than other properties. Overall only two percent of owner-occupied properties have sold. A higher share of properties that were owned by a corporation before the fire have sold--eight percent of corporate-owned properties from January 2025 have sold. The majority of these sales occurred in West Altadena and in fact, a greater share of corporate-owned properties in West Altadena have sold (11%). While property records do not specify whether these properties were owner- or renter-occupied before the fire, we assume the majority were rental units.

Recent research from SAJE has also shown how corporate investors have already exploited the Altadena community. Between February and July 2025, nearly half of parcels (47%) were purchased by corporate entities compared to only 10% bought by corporate entities in the prior year.^[https://www.saje.net/wp-content/uploads/2025/10/ConfrontingDisaster.10.10.2025.pdf] Long-term strategies to support property owners throughout their recovery are needed to help preserve opportunities for homeowners and renters in Altadena and reduce the track of speculation. 

```{r}
#### VISUAL: Sale rates by damage and area ####
df_ <- st_read(con, query="Select * from data.analysis_sales_owner_renter_sept2025") 

df <- df_ %>%
    filter(sold_after_eaton == "Sold") 

# factor level of ownership types
df$`Ownership Type`<- factor(df$owner_renter_jan, levels = c("Owner occupied, homeowner exemption",
                                                             "Likely owner occupied, no exemption",
                                                             "Renter occupied",
                                                             "Owner & renter occupied",
                                                             "Trust owned",
                                                             "Corporation owned",
                                                             "Sold to state"
                                                             ))


df <- df %>%
  select(-c(sold_after_eaton)) %>%
  mutate("% (Count)" =paste0(round(sold_prc,0),"% (", comma(sold_count)," out of ", 
                            comma(owner_renter_total), ")"),
         "Share of Housing Type Sold"=sold_prc,
         "Area" = str_to_title(area_name)) %>%
  mutate("Area" = factor(Area, levels = c("Altadena", "West", "East"))) %>%
  arrange(Area,`Ownership Type`) 


table<- df %>%
  select("Area","Ownership Type", "Share of Housing Type Sold", "% (Count)")%>%
  gt(groupname_col = "Area", rowname_col = "Owner or Renter Occupied") %>%
  tab_header(title= "Share of Properties Sold After the Fire Within Each Ownership Type")%>%
  tab_footnote(footnote=html("Catalyst California's calculations based on L.A. County Assessor Data, September 2025 and CAL FIRE Damage Data, September 2025."),placement="left") 

 table %>%
  gt_plt_bar_pct(column= 'Share of Housing Type Sold', scaled=TRUE, fill=meteorite, background="transparent")%>%
  # if needed modify column widths here 
  # cols_width(1~px(210),2~px(100),
  #            3~px(125))%>%
  opt_align_table_header("left")%>%       
   cols_align(align="left",
              columns=`Ownership Type`) %>%
  opt_table_font(font=c("Inter", default_fonts())) %>% # opt_table_font(font=c(google_font(name="Inter"),locations=NULL,default_fonts()))%>%
  tab_options(table.background.color=alabaster,
    column_labels.background.color = "transparent",
    table.font.size = px(12),
    column_labels.font.size = px(12),
    heading.align = "left",
    table.font.names = "Inter",
    heading.title.font.size = px(20),
    heading.border.bottom.color = "transparent",
    column_labels.border.bottom.width = px(3),
    column_labels.border.bottom.color = gainsboro,
    table.border.top.color = "transparent",
    table.border.bottom.color = "transparent",
    table.border.bottom.width = px(3),
    data_row.padding = px(3),
    footnotes.font.size = px(10),
    footnotes.border.bottom.color = "transparent",
    table.align='left')%>%
  tab_style(style=
            cell_borders(sides=c("top","bottom"),color=gainsboro,style="solid",weight=px(1)),locations=cells_body())%>%
   tab_style(style=
               cell_text(font=c("Inter", default_fonts()),weight=900),locations=cells_title()) # cell_text(font=c(google_font(name="Inter"),default_fonts()),weight=900),locations=cells_title())
            
```

```{r, include=FALSE}
dbDisconnect(con) 
```