---
title: ""
output: rmarkdown::html_vignette
# css: "W:\\RDA Team\\R\\cc_brandguide.css"

---

```{r, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE
)

#### Setting Up ####

library(usethis)
library(dplyr)
library(tidyverse)
library(htmltools) 
library(htmlwidgets)
library(gtExtras)
library(gt)
library(sf)
library(scales)

options(scipen=999)

source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("altadena_recovery_rebuild")

## COLORS ##
orange <- "#F25922"
peridot <- "#CEEA01"
lavender <- "#CEC4E2"
meteorite <- "#3A207D"
ccblue <- "#0860BC"

black <- "#000000"
alabaster<-"#FBFBFB"
gainsboro <- "#DEDEDE"

```

Depending on the ownership type of the property, different strategies may be need to jumpstart property owners rebuilding. Across Altadena, most significantly damaged properties are awaiting permit applications regardless of ownership type. Renter-occupied and LLC-owned properties are most likely to have not completed debris removal though these represent a small percentage of all renter-occupied or LLC-owned properties. In Altadena, over 5% of significantly damaged, renter-occupied properties have not completed debris removal.

Generally, only 1.4% of significantly damaged properties in Altadena are rebuilt across ownership types. More properties have at least applied for permits and are awaiting construction. In Altadena, 21-22% of owner-occupied and renter-occupied properties alike are awaiting the start of construction--meaning permits have been approved but construction has not started. A large percentage of all ownership types have not applied for permits. This suggests across the board property owners in Altadena are slowly progressing toward building efforts--requiring further investigation on how to tailor recovery supports for each ownership type, but in the interest of retaining the homeownership rates in Altadena, with an emphasis on owner-occupied properties.

```{r}
#### VISUAL: Permit statuses by area and owner type for significantly damaged properties  ####
df_ <- st_read(con, query="Select * from data.analysis_permits_owner_renter_jan2025") 

# pivot to wide for the table
df <- df_ %>%
  # clean up labels ahead of pivot
  mutate(rebuild_status=
           case_when(rebuild_status=="Need more info" ~ "No Evidence of Progress",
                     TRUE ~ str_to_title(rebuild_status))) %>%
  select(-count) %>% # focus on prc
    pivot_wider( # pivot
      id_cols=c(area_name,owner_renter,total), # keep total
      names_from = rebuild_status,
        values_from=prc
    )

# replace NA with 0
df[is.na(df)] <- 0

# keep principal ownership types for comparison
df <- df %>%
  filter(!owner_renter %in% c("SBE or Government owned","Church/Welfare exemption"))

df <- df %>% # 
  # rename columns for visuals
  mutate("Ownership Type"=owner_renter,
         "Total Properties"=total,
         "Area" = str_to_title(area_name)) %>%
  mutate("Area" = factor(Area, levels = c("Altadena", "West", "East"))) %>%
  # reorder, dropping no evidence of progress given the small %
 select("Area","Ownership Type", "Total Properties",
        "Debris Removal Incomplete","Awaiting Permit Application","Awaiting Construction Start", "Construction In Progress", "Rebuild Complete")%>%
  arrange(Area, desc(`Total Properties`))

table<- df %>%
  gt(groupname_col = "Area", rowname_col = "Ownership Type") %>%
  gt_plt_bar_pct(column= 'Debris Removal Incomplete',  scaled=TRUE, fill=meteorite, background=gainsboro)%>%
  gt_plt_bar_pct(column= 'Awaiting Permit Application', scaled=TRUE, fill=meteorite, background=gainsboro)%>%
  gt_plt_bar_pct(column= 'Awaiting Construction Start', scaled=TRUE, fill=meteorite, background=gainsboro)%>%
  gt_plt_bar_pct(column= 'Construction In Progress', scaled=TRUE, fill=meteorite, background=gainsboro)%>%
  gt_plt_bar_pct(column= 'Rebuild Complete', scaled=TRUE, fill=meteorite, background=gainsboro)%>%
  tab_header(title= "Rebuild Status of Significantly Damaged Properties",
             subtitle = "B y Ownership Type (as of October 2025)")%>%
  
  tab_footnote(footnote=html("Catalyst California's calculations based on L.A. County Assessor Data, January 2025 and L.A. County Public Works, October 2025."),placement="left")%>%
  # if needed modify column widths here 
  # cols_width(1~px(210),2~px(100),
  #            3~px(125))%>%
  opt_align_table_header("left")%>%
  opt_table_font(font=c("Inter", default_fonts())) %>% # opt_table_font(font=c(google_font(name="Inter"),locations=NULL,default_fonts()))%>%
  tab_options(table.background.color=alabaster,
    column_labels.background.color = "transparent",
    table.font.size = px(12),
    column_labels.font.size = px(12),
    heading.align = "left",
    table.font.names = "Inter",
    heading.title.font.size = px(20),
    heading.subtitle.font.size = px(16),
    heading.border.bottom.color = "transparent",
    column_labels.border.bottom.width = px(3),
    column_labels.border.bottom.color = gainsboro,
    table.border.top.color = "transparent",
    table.border.bottom.color = "transparent",
    table.border.bottom.width = px(3),
    data_row.padding = px(3),
    footnotes.font.size = px(10),
    footnotes.border.bottom.color = "transparent",
    table.align='left')%>%
  tab_style(style=
            cell_borders(sides=c("top","bottom"),color=gainsboro,style="solid",weight=px(1)),locations=cells_body())%>%
   tab_style(style=
               cell_text(font=c("Inter", default_fonts()),weight=900),locations=cells_title()) %>%
   tab_style(style=
               cell_text(font=c("Inter", default_fonts()),weight=600),locations=cells_title(groups="subtitle"))

table

    
```

```{r, include=FALSE}
dbDisconnect(con) 
```