---
title: ""
output: rmarkdown::html_vignette
# css: "W:\\RDA Team\\R\\cc_brandguide.css"

---

```{r, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE
)

#### Setting Up ####

library(usethis)
library(dplyr)
library(tidyverse)
library(htmltools) 
library(htmlwidgets)
library(gtExtras)
library(gt)
library(sf)
library(scales)

options(scipen=999)

source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("altadena_recovery_rebuild")

## COLORS ##
orange <- "#F25922"
peridot <- "#CEEA01"
lavender <- "#CEC4E2"
meteorite <- "#3A207D"
ccblue <- "#0860BC"

black <- "#000000"
alabaster<-"#FBFBFB"
gainsboro <- "#DEDEDE"

```

Depending on the ownership type of the property, different strategies will be needed to jumpstart rebuilding. Across Altadena, most significantly damaged properties have not had any permit applications submitted regardless of ownership type. Renter-occupied properties are generally most likely to have not completed fire debris removal though these represent a small percentage of all renter-occupied properties. Across Altadena, about 3% of significantly damaged, renter-occupied properties have not completed fire debris removal, a total of 14 properties.

Generally, less than one percent of significantly damaged properties in Altadena have completed their repairs or rebuilt across ownership types. More properties have at least applied for permits, but have not started construction. In Altadena, 27% of owner-occupied and 26% of renter-occupied properties alike have applied for and been issued permits, but have not started construction. However, a larger percentage of all ownership types have not applied for permits at all. This suggests across the board property owners in Altadena are slowly progressing toward building efforts. The few significantly damaged properties in Altadena that have been sold to state or are under other ownership are slightly further behind. About 77% of properties sold to state have not submitted permit applications and 100% of properties that are under other ownership have not submitted permit applications. Policymakers should further investigate how to tailor recovery supports for each ownership type and ensure all properties are progressing toward rebuilding. This can include holding government, corporate, or other owners accountable for rebuilding, and providing support to owner-occupied or renter-occupied properties to preserve the diversity of housing options in Altadena.

```{r}
#### VISUAL: Permit statuses by area and owner type for significantly damaged properties  ####
df_ <- st_read(con, query="Select * from data.analysis_permits_owner_renter_jan2025") 

# pivot to wide for the table
df <- df_ %>%
  # clean up labels ahead of pivot
  mutate(rebuild_status=str_to_title(rebuild_status)) %>%
  select(-count) %>% # focus on prc
    pivot_wider( # pivot
      id_cols=c(area_name,owner_renter,total), # keep total
      names_from = rebuild_status,
      values_from=prc,
      values_fill = 0
    )

# keep principal ownership types for comparison
df <- df %>%
  filter(!owner_renter %in% c("Government owned","Church or charity owned"))

# factor level of ownership types
df$`Ownership Type`<- factor(df$owner_renter, levels = c(
                                                             "Owner occupied, homeowner exemption",
                                                             "Likely owner occupied, no exemption",
                                                             "Owner & renter occupied",
                                                             "Renter occupied",
                                                             "Trust owned",
                                                             "Corporation owned",
                                                             "Sold to state",
                                                             "Other ownership"))

df <- df %>% # 
  # rename columns for visuals
  mutate(
         "Total Properties"=total,
         "Area" = str_to_title(area_name)) %>%
  mutate("Area" = factor(Area, levels = c("Altadena", "West", "East"))) %>%
  # reorder, dropping no evidence of progress given the small %
 select("Area","Ownership Type", "Total Properties",
        "Fire Debris Removal Incomplete","Permit Application Not Received","Construction Not Started", "Construction In Progress", "Repairs Or Rebuild Complete")%>%
  arrange(Area, `Ownership Type`)

table<- df %>%
  gt(groupname_col = "Area", rowname_col = "Ownership Type") %>%
  gt_plt_bar_pct(column= 'Fire Debris Removal Incomplete',  scaled=TRUE, fill=meteorite, background=gainsboro)%>%
  gt_plt_bar_pct(column= 'Permit Application Not Received', scaled=TRUE, fill=meteorite, background=gainsboro)%>%
  gt_plt_bar_pct(column= 'Construction Not Started', scaled=TRUE, fill=meteorite, background=gainsboro)%>%
  gt_plt_bar_pct(column= 'Construction In Progress', scaled=TRUE, fill=meteorite, background=gainsboro)%>%
  gt_plt_bar_pct(column= 'Repairs Or Rebuild Complete', scaled=TRUE, fill=meteorite, background=gainsboro)%>%
  tab_header(title= "Rebuild Status of Significantly Damaged Properties",
             subtitle = "By Ownership Type (as of October 2025)")%>%
  
  tab_footnote(footnote=html("Catalyst California's calculations based on L.A. County Assessor Data, January 2025 and L.A. County Public Works, October 2025."),placement="left")%>%
  # if needed modify column widths here 
  # cols_width(1~px(210),2~px(100),
  #            3~px(125))%>%
  opt_align_table_header("left")%>%
    cols_align(
    align = "left",
    columns = everything())%>%
  opt_table_font(font=c("Inter", default_fonts())) %>% # opt_table_font(font=c(google_font(name="Inter"),locations=NULL,default_fonts()))%>%
  tab_options(table.background.color=alabaster,
    column_labels.background.color = "transparent",
    table.font.size = px(12),
    column_labels.font.size = px(12),
    heading.align = "left",
    table.font.names = "Inter",
    heading.title.font.size = px(20),
    heading.subtitle.font.size = px(16),
    heading.border.bottom.color = "transparent",
    column_labels.border.bottom.width = px(3),
    column_labels.border.bottom.color = gainsboro,
    table.border.top.color = "transparent",
    table.border.bottom.color = "transparent",
    table.border.bottom.width = px(3),
    data_row.padding = px(3),
    footnotes.font.size = px(10),
    footnotes.border.bottom.color = "transparent",
    table.align='left')%>%
  tab_style(style=
            cell_borders(sides=c("top","bottom"),color=gainsboro,style="solid",weight=px(1)),locations=cells_body())%>%
   tab_style(style=
               cell_text(font=c("Inter", default_fonts()),weight=900),locations=cells_title()) %>%
   tab_style(style=
               cell_text(font=c("Inter", default_fonts()),weight=600),locations=cells_title(groups="subtitle"))

table

    
```

```{r, include=FALSE}
dbDisconnect(con) 
```