---
title: ""
output: rmarkdown::html_vignette
# css: "W:\\RDA Team\\R\\cc_brandguide.css"

---
<!-- <link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet"> -->

```{r, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE
)

## The following document outputs a visual of the housing stock before in a table format for the altadena_housing_before_and_after.html 

#### Setting Up ####

library(usethis)
library(dplyr)
library(tidyverse)
library(htmltools) 
library(htmlwidgets)
library(gtExtras)
library(gt)
library(ggplot2)
library(sf)
library(scales)
library(showtext)
library(extrafont)

options(scipen=999)

source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("altadena_recovery_rebuild")

## COLORS ##
orange <- "#F25922"
peridot <- "#CEEA01"
lavender <- "#CEC4E2"
meteorite <- "#3A207D"
ccblue <- "#0860BC"

black <- "#000000"
alabaster<-"#FBFBFB"
gainsboro <- "#DEDEDE"

## FONTS ## 

font_add_google("Inter")

# font_import()
loadfonts(device = "win")
windowsFonts()
showtext_auto()


```

While a small percentage of properties have sold, sales in Altadena surged after the Eaton fire. In April 2025, over twice as many properties sold in Altadena compared to either of the past two years. However, since April 2025, sales have dramatically dropped. This raises concerns about speculation immediately after the Eaton fire and now, possible stagnation in the housing market. This could coincide with other cooling in the L.A County market or could imply greater slowing in Altadena, which could further impact property values and raise risks for speculation.

```{r}
#### VISUAL: Sale rates by damage and area ####
df_ <- st_read(con, query="Select * from data.analysis_sales_month_sept2025") 

df <- df_ %>%
  mutate(month=case_when(last_sale_month=="01" ~ "Jan",
                         last_sale_month=="02" ~ "Feb",
                         last_sale_month=="03" ~ "Mar",
                         last_sale_month=="04" ~ "Apr",
                         last_sale_month=="05" ~ "May",
                         last_sale_month=="06" ~ "Jun",
                         last_sale_month=="07" ~ "Jul",
                         last_sale_month=="08" ~ "Aug",
                         TRUE ~ NA
  ))

df$month <- factor(df$month, levels = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug"))
                   
    ggplot(df, aes(x = month, y = sales_count, color = last_sale_year, group = last_sale_year)) +
      geom_line() +
      geom_point() +
      labs(title = "Altadena Residential Sales by Year and Month",
           x = "",
           y = "Properties Sold Count",
           caption="Catalyst California's calculations of L.A. County Assessor Data, September 2025.") +
      theme_minimal() + # Optional: for a cleaner theme
      labs(color="") +
      theme(legend.position="top",
             panel.grid.minor.x = element_blank(),
  # panel.grid.minor.y = element_blank(),
plot.title = element_text(family = "Inter", size=14, face="bold"),
        # plot.subtitle = element_text(family = "Inter", size = 12, color = black),
        plot.caption = element_text(family = "Inter", size = 9, color = "#6F6F6F", hjust=0),
        legend.title = element_text(family = "Inter", size = 12,),
        legend.text = element_text(family = "Inter", size = 12),
axis.text.y=element_text(family = "Inter", size = 10, color = "#6F6F6F"),
  axis.text.x=element_text(family = "Inter", size = 11, color="black"),
axis.title.y=element_text(family = "Inter", size = 10, color = "#6F6F6F"),
axis.title.x=element_text(family = "Inter", size = 10, color = "#6F6F6F")
  ) +
      scale_color_manual(values=c(lavender, meteorite, orange))
    
            
```

```{r, include=FALSE}
dbDisconnect(con) 
```