---
title: 'Preserving Altadena Housing in Recovery and Rebuilding'
subtitle: "Changes in Housing Stock Before and After the Eaton Fire"
author: "Catalyst California"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 3
    toc_float: yes
---
<!-- link to Google Font -->
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE
)

# Library and environment set up ----

library(sf)
library(dplyr)
# library(highcharter)
library(rmapshaper)
library(leaflet)
library(htmlwidgets)
# library(tigris)
# library(tidycensus)
library(stringr)
library(showtext)
library(extrafont)
library(readxl)
library(tidyverse)
library(janitor)
library(flextable)
library(kableExtra)
# library(ggplot2)
# options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))
options(scipen=999)

# Colors and Fonts ----
## COLORS ##
papaya <- "#F25922"
peridot <- "#CEEA01"
lavender <- "#CEC4E2"
meteorite <- "#3A207D"
ccblue <- "#0860BC"

black <- "#000000"
alabaster<-"#FBFBFB"
gainsboro <- "#DEDEDE"

## FONTS ## 

font_add_google("Inter")

# font_import()
loadfonts(device = "win")
windowsFonts()
showtext_auto()

# Load in and prep shapes ----
source("W:\\RDA Team\\R\\credentials_source.R")

con_alt <- connect_to_db("altadena_recovery_rebuild")

```

# Introduction
EMG - what this page covers

# Prior Research
MTK - bullets on findings from literature review

# Key Findings
EMG

# Methods
EMG


# Altadena Before

## Map

```{r map-before, child='Embeds/map_before_eaton.Rmd', echo=FALSE}
```


## Residential Housing Types

```{r restype-before, child='Embeds/visual_housing_before.Rmd',echo=FALSE}
```

## Rental Units

```{r units-before, child='Embeds/visual_units_before.Rmd', echo=FALSE}
```

## Ownership Types

```{r owner-before, child='Embeds/visual_homeowner_before.Rmd', echo=FALSE}
```


## Average Square Footage

```{r sqfeet-before, child='Embeds/visual_sqftg_before.Rmd', echo=FALSE}
```

# Altadena After

## Map
```{r map-after, child='Embeds/map_after_eaton.Rmd', echo=FALSE}
```


## Residential Types Lost

```{r restype-after, child='Embeds/visual_housing_after.Rmd', echo=FALSE}
```

```{r multi-after, child='Embeds/visual_multifamily_after.Rmd', echo=FALSE}
```

## Rental Units Lost

```{r units-after, child='Embeds/visual_units_after.Rmd', echo=FALSE}
```

## Ownership Types Lost

```{r owner-after, child='Embeds/visual_homeowner_after.Rmd', echo=FALSE}
```

## Average Square Footage Lost

```{r sqftg-after, child='Embeds/visual_sqftg_after.Rmd', echo=FALSE}
```

# Altadena Rebuilding

## Map
```{r planning-map, child='Embeds/map_damage.Rmd', echo=FALSE}
```

## Properties Sold

```{r sold-month, child='Embeds/visual_sales_bymonth.Rmd', echo=FALSE}
```

```{r sold-damage, child='Embeds/visual_sales_damage.Rmd', echo=FALSE}
```

```{r sold-owner, child='Embeds/visual_sales_owner_renter.Rmd', echo=FALSE}
```

## Permits

```{r permits-area, child='Embeds/visual_permits_damage.Rmd', echo=FALSE}
```

```{r permits-owner, child='Embeds/visual_permits_owner.Rmd', echo=FALSE}
```

## Lead Contamination 

```{r res-lead, child='Embeds/visual_restype_highlead.Rmd', echo=FALSE}
```

```{r owner-lead, child='Embeds/visual_homeowner_highlead.Rmd', echo=FALSE}
```

## Fire Hazards

```{r fhsz, child='Embeds/visual_damage_fhsz.Rmd', echo=FALSE}
```

## Zoning

```{r zoning, child='Embeds/visual_zoning.Rmd', echo=FALSE}
```

# Recommendations

# Conclusion