---
title: 'Altadena Before and After the Eaton Fire'
subtitle: "Changes in Altadena Housing Stock Post the Eaton Fire"
author: "Catalyst California"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 2
    toc_float: yes
---
<!-- link to Google Font -->
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE
)

# Library and environment set up ----

library(sf)
library(dplyr)
# library(highcharter)
library(rmapshaper)
library(leaflet)
library(htmlwidgets)
# library(tigris)
# library(tidycensus)
library(stringr)
library(showtext)
library(extrafont)
library(readxl)
library(tidyverse)
library(janitor)
library(flextable)
library(kableExtra)
# library(ggplot2)
# options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))
options(scipen=999)

# Colors and Fonts ----
## COLORS ##
papaya <- "#F25922"
peridot <- "#CEEA01"
lavender <- "#CEC4E2"
meteorite <- "#3A207D"
ccblue <- "#0860BC"

black <- "#000000"
alabaster<-"#FBFBFB"
gainsboro <- "#DEDEDE"

## FONTS ## 

font_add_google("Inter")

# font_import()
loadfonts(device = "win")
windowsFonts()
showtext_auto()

# Load in and prep shapes ----
source("W:\\RDA Team\\R\\credentials_source.R")

con_alt <- connect_to_db("altadena_recovery_rebuild")

```

# Introduction
EMG - what this page covers

# Prior Research
MTK - bullets on findings from literature review

# Key Findings
EMG

# Housing Stock Before the Fire -- MTK

```{r}
# map of all properties in East/West Altadena with layers for residential housing type, homeowner/renter, unit size, and square footage
```

## Residential Housing Types

```{r restype-before, child='Embeds/visual_housing_before.Rmd',echo=FALSE}
```

## Homeowner and Rental Properties
Distribution of homeowner and rental properties prior to the fire in Altadena, East Altadena, West Altadena
```{r owner-before, child='Embeds/visual_homeowner_before.Rmd', echo=FALSE}
```

## Residential Unit Types
Total and types of units prior to the fire in Altadena, East Altadena, West Altadena
```{r units-before, child='Embeds/visual_units_before.Rmd', echo=FALSE}
```

## Square Footage
Average square footage of residential properties prior to the fire in Altadena, East Altadena, West Altadena
```{r sqfeet-before, child='Embeds/visual_sqftg_before.Rmd', echo=FALSE}
```

# Characteristics of Properties Significantly Damaged -- MTK

```{r}
# map of properties significant damaged or destroyed with layers for residential housing type, homeowner/renter, total units, square footage

```

## Types of Residential Housing Types Lost
Distribution of housing types destroyed or damaged in Altadena, East, West compared to original percentages
```{r restype-after, child='Embeds/visual_housing_after.Rmd', echo=FALSE}
```

## Homeowner and Rental Properties Lost
Distribution of homeowner and rental properties destroyed or damaged in Altadena, East, West compared to original percentages
```{r owner-after, child='Embeds/visual_homeowner_after.Rmd', echo=FALSE}
```

## Residential Units Lost
Total and types of units destroyed or damaged in Altadena, East, West
```{r units-after, child='Embeds/visual_units_after.Rmd', echo=FALSE}
```

## Average Square Footage Lost
Average square footage of residential properties destroyed or damaged in Altadena, East, West compared to original percentages
```{r sqftg-after, child='Embeds/visual_sqftg_after.Rmd', echo=FALSE}
```

# Changes Since the Fire -- EMG or JZ

```{r}
# map of properties being sold or rebuilt

```


## Properties Sold
How many properties have sold in Altadena, East, and West.--what are the characteristics of those sold
```{r}
```

## Properties Being Rebuilt
How many properties that have been damaged or rebuilt are in the rebuilding phases--what are the characteristics of those being rebuilt
```{r}
```


# Environmental and Zoning Considerations -- EMG

```{r}
# map with damaged/destroyed properties and then lead, zoning, and fire hazard zones

```

## Lead Contamination 
Share of damaged properties in high lead areas in Altadena, East, and West
```{r}
```

## Zoning
What are the zoning uses of properties in Altadena, East, and West
```{r}
```

## Fire Hazards
How many properties are in high fire zones
```{r}
```

# Recommendations

# Conclusion