---
title: 'Toward an Equitable Recovery After Disaster - DRAFT'
subtitle: "Takeaways From Changes in Altadena Housing Before and After the Eaton Fire"
author: "Catalyst California"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 3
    toc_float: yes
---
<!-- link to Google Font -->
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE
)

# Library and environment set up ----

library(sf)
library(dplyr)
# library(highcharter)
library(rmapshaper)
library(leaflet)
library(htmlwidgets)
# library(tigris)
# library(tidycensus)
library(stringr)
library(showtext)
library(extrafont)
library(readxl)
library(tidyverse)
library(janitor)
library(flextable)
library(kableExtra)
# library(ggplot2)
# options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))
options(scipen=999)

# Colors and Fonts ----
## COLORS ##
papaya <- "#F25922"
peridot <- "#CEEA01"
lavender <- "#CEC4E2"
meteorite <- "#3A207D"
ccblue <- "#0860BC"

black <- "#000000"
alabaster<-"#FBFBFB"
gainsboro <- "#DEDEDE"

## FONTS ## 

font_add_google("Inter")

# font_import()
loadfonts(device = "win")
windowsFonts()
showtext_auto()

# Load in and prep shapes ----
source("W:\\RDA Team\\R\\credentials_source.R")

con_alt <- connect_to_db("altadena_recovery_rebuild")

```

# Introduction 

Altadena is rich with historical significance and threads of community resilience. Altadena is an unincorporated region of Los Angeles County below the San Gabriel Mountains and to the north of the City of Pasadena. Throughout the years, the region has become a vibrant hub for communities of color to call home. When housing was racially segregated in the early to mid 1900s, Altadena became one of the few places where Black families could own property. The community grew and became a place where communities of color could thrive. Earlier this year, the Eaton fire devastated Altadena and now the community faces challenges in recovery and rebuilding this essential part of greater Los Angeles.

A network of resident coalitions are working together to rebuild and retain the heart of the Altadena community before the fires, and funders are identifying opportunities to support their efforts. In order to reach an equitable recovery, investments and policies meant to support Altadenans must understand the makeup of the community's housing, owners, and tenants before the fires. Additionally, progress toward rebuilding, and actions undermining recovery, must be tracked to identify where property owners are succeeding, falling behind, or in need of additional support.

Toward that end, this report aims to outline (1) the composition of Altadena's housing stock before the fire, (2) the housing stock destroyed in the fire, and (3) progress and challenges toward rebuilding, including risks to a safe and equitable recovery. Given the disproportionate impact of the fire on West Altadena, we detail differences between West and East Altadena in their housing stock before and after the Eaton Fire. Each section includes detailed maps and assessments outlining how single-family homes, multifamily homes, condominiums, tenants, homeowners, and landlords were affected by the fire. We hope this information and prior research can help advocates, decision-makers, and funders in their pursuit to rebuild Altadena from the ashes for all Altadenans. 

## Prior Research

Between January 7th and 31st, 2025, the Eaton Fire tore through the Altadena community. Numerous research studies find that Black families and communities of color faced higher loss than others. For example, Black individuals and families experienced severe damage and destruction of their homes at a rate of almost 60%, the highest rate compared to other race/ethnic groups.^[UCLA. (2025, October 9). UCLA analysis finds Altadena faces uneven wildfire recovery as Black, Latino, and AAPI homeowners struggle to rebuild. Pasadena Now. https://pasadenanow.com/main/ucla-analysis-finds-altadena-faces-uneven-wildfire-recovery-as-black-latino-and-aapi-homeowners-struggle-to-rebuild ] Compounding systemic issues makes recovery even more difficult for Black families and communities of color. From Black homeowners in Altadena, 28% report being severely cost-burdened, where only 13% of non-Black homeowners report being cost-burdened.^[UCLA Bunche Center. (n.d.). LA wildfires: Impacts on Altadena’s Black community (P. Ong, C. Pech, L. Frasure, S. Comandur, E. Lee, & S. R. González). Retrieved from https://bunchecenter.ucla.edu/wildfires-altadena-black-community/] Furthermore, rising housing costs and tenure decline leave a vulnerable population to be displaced and more likely to risk losing housing and a loss of generational wealth.^[SAJE (Strategic Actions for a Just Economy). (2025, May). Compounding disaster: A spatial analysis of housing risk and speculation in post-fire Altadena.]  

It is imperative that policymakers understand these risks and disparities and address systemic issues that can compound inequities after the dust of the debris settles. Unfortunately, past disasters find that recovery efforts often fail to reach the most vulnerable populations.^[Ney, J., & Palmer, D. (2025, July 30). Why disaster aid fails the most vulnerable. American Inequality. https://americaninequality.substack.com/p/why-disaster-aid-fails-the-most-vulnerable-givedirectly] When we continue to reinforce inequities in disaster recovery, communities are subjected to challenges far worse than the initial devastation.  

Less than a year following the Eaton fire, the data already shows that county and state leaders are not doing enough. Prior research finds that 70% of severely damaged homes show no signs of action toward rebuilding and the homes that have been resold, almost two-thirds have been given to outside investors.^[UCLA. (2025, October 9). UCLA analysis finds Altadena faces uneven wildfire recovery as Black, Latino, and AAPI homeowners struggle to rebuild. Pasadena Now. https://pasadenanow.com/main/ucla-analysis-finds-altadena-faces-uneven-wildfire-recovery-as-black-latino-and-aapi-homeowners-struggle-to-rebuild ] Many community members and longtime residents are concerned that local ownership has been lost in a county already facing a lack of affordable housing and some of the highest rent costs in the country.^[Phillips, S. (2025, February). The Palisades and Eaton fires: Neighborhood data and potential housing market effects. UCLA Lewis Center for Regional Policy Studies. https://www.lewis.ucla.edu/research/the-palisades-and-eaton-fires-neighborhood-data-and-potential-housing-market-effects/] Another ripple effect of the fires and other climate related disasters are the hike of insurance costs, another force increasing the already high housing cost burden.^[U.S. Department of the Treasury. (2025, January 16). Homeowners insurance costs rising, availability declining as climate-related events take their toll. https://home.treasury.gov/news/press-releases/jy2791]  

When the Eaton fire hit this region earlier this year, the devastation not only destroyed properties but one of the most historically unique regions in Los Angeles. While properties can be built again with cement and gravel, it is going to take equity and community driven work to rebuild the haven Altadena was for so many Black families and communities of color in the county.

# Key Findings

To support an equitable recovery where Altadenans are able to return to and rebuild in their community, we emphasize the following key findings from our analysis: 

**Before**

* In January 2025, immediately before the Eaton Fire, Altadena was largely a low density, high owner-occupied community. Nearly 91% of residential properties were single-family homes and over 50% of properties were owner-occupied.

* Multifamily and renter-occupied properties existed in Altadena at low rates before the fire, less than ten and fifteen percent of properties respectively.

**After**

* The Eaton Fire caused greater devastation and property loss in West Altadena. Nearly half of multifamily properties in West Altadena were significantly damaged and over half of single-family homes and condominiums were.

* Nearly half of rental units in West Altadena were significantly damaged exceeding damage rates in East Altadena. Similarly, over half of owner-occupied units in West Altadena were significantly damaged.

**Rebuilding**

* Altadena residential properties surged in sales in April 2025 following the fire. That surge has now slowed, and overall a small percentage of properties have sold.

* Significantly damaged properties make up a majority of the sales compared to properties that sustained no damage, and most properties sold were in West Altadena.

* Across the board, property owners in Altadena are stuck in rebuilding efforts. Only one in three significantly damaged properties in Altadena have at least applied for a permit to begin construction, meaning the greater majority of properties are stuck in a stage after debris removal with no permits to rebuild.

* Across Altadena, significantly damaged multifamily properties and condominiums are more likely to be in high-lead level areas compared to single-family homes.

* Across Altadena, even properties that sustained no damage are in high lead areas--meaning that additional environmental testing and mitigation should have included no damaged properties as well as damaged properties.

* While West Altadena suffered more damage in th fire, a lower percentage of significantly damaged properties in the area are considered to be within a Very High Fire Hazard Zone compared to East Altadena--raising questions about how property owners in Altadena, and other parts of California, can prepare for human- and climate-caused disasters like the Eaton fire without adequate hazard measures.

# Methods 

While the Eaton Fire affected the Altadena and Pasadena communities, we focused our analysis on Altadena community boundaries in order to gather a full assessment of what this defined community looked like before and after the fire. This also allowed us to focus on scraping permit data for unincorporated L.A. County. For our boundaries, we split Altadena into West and East Altadena using the unincorporated community/place boundaries and splitting the community by West and East of Lake Avenue.

To analyze property characteristics before the fire, we gathered L.A. County Assessor data from January 2025. We focused on residential properties in Altadena, or parcels with a residential use code and that were not recorded as vacant land. While we initially included mixed use properties, we found these properties to comprise a very small share of Altadena properties. In January 2025, only 20 properties were mixed use properties with residential units.

To analyze housing lost in the fire, we utilized data from the CalFire Damage Inspection database and matched these structure-level records to their corresponding parcel records. This also required creating a typology to classify parcels as mostly damaged or destroyed even when they had multiple structures with different damage characteristics. We classified parcels into these categories to help focus on residential properties that sustained greater damage and make our findings more easily interpretable (by including fewer categories):

* No Damage: Parcels that had no damage recorded on any structure on the property, or were inaccessible.
* Some Damage: Parcels that included at least one structure that was affected (1-9% destroyed) and/or had minor damage (10-25% destroyed) alone or in combination with structures that had no damage.
* Significant Damage or Destroyed: Parcels that included at least one structure that had major damage (26-50% destroyed) or destroyed (>50% destroyed) alone or in combination with another damage category. We count these properties as "lost" residential properties.

To analyze progress toward rebuilding, we gathered L.A. County Assessor Data from September 2025--our source for sales data and property characteristics. We also scraped permit data from L.A. County for October 2025 as well as debris removal data from the Army Corps of Engineers to track progress toward cleanup and rebuilding.

# Altadena Before

Understanding Altadena's housing composition before the Eaton Fire can help track progress toward an equitable recovery and rebuilding. We analyzed the residential housing types, ownership characteristics, rental units, and square footage of properties before Eaton Fire based on L.A. County Assessor data from January 2025.

## Map

In January 2025, immediately before the Eaton Fire, Altadena was primarily a low-density residential community with few multifamily units or mixed use properties. Higher density multifamily properties were concentrated around Lake and Allen Avenue and the southern blocks of West and East Altadena. Most properties were owner or renter-occupied.

```{r map-before, child='Embeds/map_before_eaton.Rmd', echo=FALSE}
```


## Residential Housing Types

```{r restype-before, child='Embeds/visual_housing_before.Rmd',echo=FALSE}
```

## Ownership Types

```{r owner-before, child='Embeds/visual_homeowner_before.Rmd', echo=FALSE}
```

## Rental Units

```{r units-before, child='Embeds/visual_units_before.Rmd', echo=FALSE}
```

## Average Square Footage

```{r sqfeet-before, child='Embeds/visual_sqftg_before.Rmd', echo=FALSE}
```

# Altadena After

The Eaton Fire burned through the Altadena community for nearly 24 days. Prior research documents how West Altadena was disproportionately devastated by the fire with all fire-related deaths occurring West of Lake Avenue.^[https://www.nbcnews.com/news/us-news/eaton-fire-deaths-los-angeles-evacuation-orders-took-hours-rcna188729] Due to systemic failures in alert and evacuation operations, West Altadena did not receive evacuation orders until nine hours after the Eaton Fire began. A change in wind speed and direction aggressively pushed the fire west early in the morning of January 8th.^[https://file.lacounty.gov/SDSInter/bos/supdocs/207915.pdf] 

We analyzed the destruction from the fire West and East of Lake Avenue examining the share of housing types significantly damaged by the fire. These trends help inform the housing stock lost to the fire and in need of rebuilding in each community; and therefore, also the Altadenans most at risk of displacement.

## Map

West Altadena was largely destroyed by the fire with the fire perimeter covering a significant area of the community. There are pockets of housing that sustained no damage, such as portions along Glenrose Avenue and blocks around Lake Avenue and Dolores Drive. East Altadena had more pockets with no damage such as along North Hollister Avenue, East Mendocino Street, and East Calaveras Street. However, throughout Altadena, even pockets with no damage are surrounded by properties significantly damaged or destroyed. 

```{r map-after, child='Embeds/map_after_eaton.Rmd', echo=FALSE}
```


## Residential Types Lost

```{r restype-after, child='Embeds/visual_housing_after.Rmd', echo=FALSE}
```

```{r multi-after, child='Embeds/visual_multifamily_after.Rmd', echo=FALSE}
```

## Ownership Types Lost

```{r owner-after, child='Embeds/visual_homeowner_after.Rmd', echo=FALSE}
```

## Rental Units Lost

```{r units-after, child='Embeds/visual_units_after.Rmd', echo=FALSE}
```

## Average Square Footage Lost

```{r sqftg-after, child='Embeds/visual_sqftg_after.Rmd', echo=FALSE}
```

# Altadena Rebuilding

The dust from the fire has settled, but Altadena faces a long recovery ahead. To rebuild, varying types of property owners need support in navigating insurance, permits, hazards, and new building standards. Meanwhile, predatory investors have already begun displacing residents and taken advantage of owners when they were most vulnerable. We analyzed the properties that have sold and those that are in progress toward rebuilding as well as the hazard and zoning considerations owners must take into account when rebuilding. All of these factors will affect the path and result of Altadena recovering. Based on September 2025 data from the L.A. County Assessor's office, we mapped and analyzed the properties that have sold. Based on L.A. County permit data we scraped in October 2025, we analyzed the share of significantly damaged properties that are on the road to rebuilding.

## Map

Few properties have sold since the fire, but most are in West Altadena with several homes selling around Lake Avenue and Wapello Street. Meanwhile, most properties have completed debris removal but have not sought an initial permit to rebuild. As properties rebuild, they must also contend with fire and lead hazards in the community. The northern portions of West and East Altadena are in fire hazard zones, with the northern most portions in very high hazard areas that require additional defense measures. Meanwhile, the southern and most western portions of Altadena face higher lead levels--downwind from the fire. Beyond the direct fire perimeter, homes outside of the Eaton Fire burn area are in grids with average lead levels above recommended screening thresholds.

```{r planning-map, child='Embeds/map_damage.Rmd', echo=FALSE}
```

## Properties Sold

```{r sold-month, child='Embeds/visual_sales_bymonth.Rmd', echo=FALSE}
```

```{r sold-damage, child='Embeds/visual_sales_damage.Rmd', echo=FALSE}
```

```{r sold-owner, child='Embeds/visual_sales_owner_renter.Rmd', echo=FALSE}
```

## Permits

```{r permits-area, child='Embeds/visual_permits_damage.Rmd', echo=FALSE}
```

```{r permits-owner, child='Embeds/visual_permits_owner.Rmd', echo=FALSE}
```

## Lead Contamination 

```{r res-lead, child='Embeds/visual_restype_highlead.Rmd', echo=FALSE}
```

```{r owner-lead, child='Embeds/visual_homeowner_highlead.Rmd', echo=FALSE}
```

## Fire Hazards

```{r fhsz, child='Embeds/visual_damage_fhsz.Rmd', echo=FALSE}
```

## Zoning

```{r zoning, child='Embeds/visual_zoning.Rmd', echo=FALSE}
```

# Recommendations - in progress

* Strategies to reduce speculation--buying up of properties
* Strategies to support rebuilding that are tailored to property and ownership types, e.g., what does it look like to support single-family homes in rebuilding vs. multifamily and homeowners and renters
* Help safeguard multfamily properties and rental units to prevent speculation and displacement of Altadenan renters, emphasis on recovering multifamily and rental units in West Altadena that were disproprtionately affected
* Provide tailored supports to homeowners seeking to rebuild to ensure homeownership remains an opportunity for existing community members and prevent further speculation
* Track SB 9 permits to ensure preservation of Altadena's single family homes while also ensuring multifamily housing is available
* Track property sales and values in Altadena to protect against speculation but also home value as a source of wealth for Altadena community members
* Identify and address hurdles to rebuilding for Altadena community members including at key stages in permitting process--compiling necessary plans and paperwork to obtain permit and then begin construction
* provide additional mitigation supports to homes in high lead testing grids particularly rental units, multifamily properties in East Altadena, and single family homes in West Altadena. This includes both intact and significantly damaged homes.
* provide support for Altadenan community members to safely rebuild complying with defense space requirements 
* pursue broader disaster preparedness across communities given experience of Altadena community that was mostly not in a very high fire zone yet endured one of the worst human and climate caused catastrophes in California in the last decade


# Conclusion - in progress