---
title: 'Toward an Equitable Recovery After Disaster - DRAFT'
subtitle: "Takeaways From Changes in Altadena Housing Before and After the Eaton Fire"
author: "Catalyst California"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 3
    toc_float: yes
---
<!-- link to Google Font -->
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE
)

# Library and environment set up ----

library(sf)
library(dplyr)
# library(highcharter)
library(rmapshaper)
library(leaflet)
library(htmlwidgets)
# library(tigris)
# library(tidycensus)
library(stringr)
library(showtext)
library(extrafont)
library(readxl)
library(tidyverse)
library(janitor)
library(flextable)
library(kableExtra)
# library(ggplot2)
# options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))
options(scipen=999)

# Colors and Fonts ----
## COLORS ##
papaya <- "#F25922"
peridot <- "#CEEA01"
lavender <- "#CEC4E2"
meteorite <- "#3A207D"
ccblue <- "#0860BC"

black <- "#000000"
alabaster<-"#FBFBFB"
gainsboro <- "#DEDEDE"

## FONTS ## 

font_add_google("Inter")

# font_import()
loadfonts(device = "win")
windowsFonts()
showtext_auto()

# Load in and prep shapes ----
source("W:\\RDA Team\\R\\credentials_source.R")

con_alt <- connect_to_db("altadena_recovery_rebuild")

```

# Introduction 

Altadena is rich with historical significance and threads of community resilience. Altadena is an unincorporated region of Los Angeles County below the San Gabriel Mountains and to the north of the City of Pasadena. Throughout the years, the region has become a vibrant hub for communities of color to call home. When housing was racially segregated in the early to mid 1900s, Altadena became one of the few places where Black families could own property. The community grew and became a place where communities of color could thrive. Earlier this year, the Eaton Fire devastated Altadena and now the community faces challenges in recovery and rebuilding this essential part of greater Los Angeles.

A network of resident coalitions are working together to rebuild and retain the heart of the Altadena community before the fires, and funders are identifying opportunities to support their efforts. In order to reach an equitable recovery, investments and policies meant to support Altadenans must understand the makeup of the community's housing, owners, and tenants before the fires. Additionally, progress toward rebuilding, and actions undermining recovery, must be tracked to identify where property owners are succeeding, falling behind, or in need of additional support.

Toward that end, this report aims to outline: (1) the composition of Altadena's housing stock before the fire, (2) the housing stock destroyed in the fire, and (3) progress and challenges toward rebuilding, including risks to a safe and equitable recovery. Given the disproportionate impact of the fire on West Altadena, we detail differences between West and East Altadena in their housing stock before and after the Eaton Fire. Each section includes detailed maps and assessments outlining how single-family homes, multifamily homes, condominiums, tenants, homeowners, and landlords were affected by the fire. We hope this information and prior research can help advocates, decision-makers, and funders in their pursuit to rebuild Altadena from the ashes for all Altadenans. 

## Prior Research

Between January 7th and 31st, 2025, the Eaton Fire tore through the Altadena community. Numerous research studies find that Black families and communities of color faced higher loss than others. For example, Black individuals and families experienced severe damage and destruction of their homes at a rate of almost 60%, the highest rate compared to other race/ethnic groups.^[UCLA. https://latino.ucla.edu/research/who-is-coming-home-altadena/] Compounding systemic issues makes recovery even more difficult for Black families and communities of color. From Black homeowners in Altadena, 28% report being severely cost-burdened, where only 13% of non-Black homeowners report being cost-burdened.^[UCLA Bunche Center. (n.d.). LA wildfires: Impacts on Altadena’s Black community (P. Ong, C. Pech, L. Frasure, S. Comandur, E. Lee, & S. R. González). Retrieved from https://bunchecenter.ucla.edu/wildfires-altadena-black-community/] Furthermore, rising housing costs and tenure decline leave a vulnerable population to be displaced and more likely to risk losing housing and a loss of generational wealth.^[SAJE (Strategic Actions for a Just Economy). (2025, May). Compounding disaster: A spatial analysis of housing risk and speculation in post-fire Altadena. Retrieved from https://www.saje.net/wp-content/uploads/2025/05/compounding-disaster-FINAL_WEB.pdf.]  

It is imperative that policymakers understand these risks and disparities and address systemic issues that can compound inequities after the dust of the debris settles. Unfortunately, past disasters find that recovery efforts often fail to reach the most vulnerable populations such as lower-income families that studies found needed almost three times as much time to financially recover from the loss.^[Ney, J., & Palmer, D. (2025, July 30). Why disaster aid fails the most vulnerable. American Inequality. https://americaninequality.substack.com/p/why-disaster-aid-fails-the-most-vulnerable-givedirectly] One of the reasons these vulnerable communities are impacted more is because of pre-existing systemic issues such as lower income communities of color living in areas with a higher likelihood of climate disasters, yet with a lower likelihood of being fully insured. When we continue to reinforce inequities in disaster recovery, communities are subjected to challenges far worse than the initial devastation.  

Less than a year following the Eaton Fire, the data already shows that county and state leaders are not doing enough. Prior research finds that 70% of severely damaged homes show no signs of action toward rebuilding and of the homes that have been resold, almost two-thirds have been given to outside investors.^[UCLA. (2025, October 9). UCLA analysis finds Altadena faces uneven wildfire recovery as Black, Latino, and AAPI homeowners struggle to rebuild. Pasadena Now. https://pasadenanow.com/main/ucla-analysis-finds-altadena-faces-uneven-wildfire-recovery-as-black-latino-and-aapi-homeowners-struggle-to-rebuild ] Many community members and longtime residents are concerned that local ownership has been lost in a county already facing a lack of affordable housing and some of the highest rent costs in the country.^[Phillips, S. (2025, February). The Palisades and Eaton Fires: Neighborhood data and potential housing market effects. UCLA Lewis Center for Regional Policy Studies. https://www.lewis.ucla.edu/research/the-palisades-and-eaton-fires-neighborhood-data-and-potential-housing-market-effects/] Another ripple effect of the fires and other climate related disasters are the hike of insurance costs, another force increasing the already high housing cost burden.^[U.S. Department of the Treasury. (2025, January 16). Homeowners insurance costs rising, availability declining as climate-related events take their toll. https://home.treasury.gov/news/press-releases/jy2791] Insurance companies are also fleeing California altogether because of the negative cost-benefit impact on their bottom line--making it no longer lucrative for them to operate in the state. In turn, the California FAIR Plan, the state-chartered insurer, is having to shoulder insurance coverage costs for formerly private policyholders, posing yet another barrier to recovery.^[https://www.eenews.net/articles/insurers-flee-wildfire-prone-california-despite-state-assistance/]

When the Eaton Fire hit this region earlier this year, the devastation not only destroyed properties but one of the most historically unique regions in Los Angeles. While properties can be built again with cement and gravel, it is going to take equity and community driven work to rebuild the haven Altadena was for so many Black families and communities of color in the county.

# Key Findings

To support an equitable recovery where Altadenans are able to return to and rebuild in their community, we emphasize the following key findings from our analysis: 

**Before**

* In January 2025, immediately before the Eaton Fire, Altadena was largely a low density, high owner-occupied community. Nearly 91% of residential properties were single-family homes and over 50% of properties were owner-occupied.

* Multifamily and renter-occupied properties existed in Altadena at low rates before the fire, less than ten and fifteen percent of properties respectively.

**After**

* The Eaton Fire caused greater devastation and property loss in West Altadena. Nearly half of multifamily properties in West Altadena were significantly damaged and over half of single-family homes and condominiums were.

* Nearly half of rental units in West Altadena were significantly damaged exceeding damage rates in East Altadena. Similarly, over half of owner-occupied units in West Altadena were significantly damaged.

**Rebuilding**

* Altadena residential properties surged in sales in April 2025 following the fire. That surge has now slowed, and overall a small percentage of properties have sold.

* Significantly damaged properties make up a majority of the sales compared to properties that sustained no damage, and most properties sold were in West Altadena.

* Across the board, property owners in Altadena are stuck in rebuilding efforts. Only one in three significantly damaged properties in Altadena have at least applied for a permit to begin construction, meaning the great majority of properties are stuck in a stage after debris removal with no permits to rebuild.

* Across Altadena, nearly half of significantly damaged, single-family homes are in high-lead level areas; and even higher shares of significantly damaged, multifamily properties and condominiums are in high-lead level areas.

* Across Altadena, even properties that sustained no damage are in high lead areas--meaning that additional environmental testing and mitigation should have included no damaged properties as well as damaged properties.

* Although 71% of significantly damaged properties in Altadena occurred in West Altadena, most of these properties are not considered to be in a Fire Hazard Zone--raising questions about how property owners here, and across California, can prepare for future climate disasters without adequate hazard measures.

# Methods 

While the Eaton Fire affected the Altadena and Pasadena communities, we focused our analysis on Altadena community boundaries in order to gather a full assessment of what this defined community looked like before and after the fire. This also allowed us to focus on scraping permit data for unincorporated L.A. County. For our boundaries, we split Altadena into West and East Altadena using the unincorporated community/place boundaries and splitting the community by West and East of Lake Avenue.

To analyze property characteristics before the fire, we gathered L.A. County Assessor data from January 2025. We focused on residential properties in Altadena, or parcels with a residential use code that were not recorded as vacant land. While we initially included mixed use properties, we found these properties to comprise a very small share of Altadena properties. In January 2025, only 20 properties were mixed use properties with residential units.

To analyze housing lost in the fire, we utilized data from the CalFire Damage Inspection database and matched these structure-level records to their corresponding parcel records. This also required creating a typology to classify parcels as mostly damaged or destroyed even when they had multiple structures with different damage characteristics. We classified parcels into these categories to help focus on residential properties that sustained greater damage and make our findings more easily interpretable (by including fewer categories):

* No Damage: Parcels that had no damage recorded on any structure on the property, or were inaccessible.

* Some Damage: Parcels that included at least one structure that was affected (1-9% destroyed) and/or had minor damage (10-25% destroyed) alone or in combination with structures that had no damage.

* Significant Damage or Destroyed: Parcels that included at least one structure that had major damage (26-50% destroyed) or destroyed (>50% destroyed) alone or in combination with another damage category. We count these properties as "lost" residential properties.

To track progress toward cleanup and rebuilding, we gathered L.A. County Assessor Data from September 2025--our source for sales data and property characteristics. We also scraped permit data from L.A. County's Electronic Permitting and Inspections portal (EPIC-LA) for October 2025 as well as wildfire debris removal data from the U.S. Army Corps of Engineers.

# Altadena Before

Understanding Altadena's housing composition before the Eaton Fire can help track progress toward an equitable recovery and rebuilding. We analyzed the residential housing types, ownership characteristics, rental units, and square footage of properties before Eaton Fire based on L.A. County Assessor data from January 2025.

## Map

In January 2025, immediately before the Eaton Fire, Altadena was primarily a low-density residential community with few multifamily units or mixed use properties. Higher density multifamily properties were concentrated around Lake and Allen Avenue and the southern blocks of West and East Altadena. The majority of properties were owner-occupied but many were also renter-occupied..

```{r map-before, child='Embeds/map_before_eaton.Rmd', echo=FALSE}
```


## Residential Housing Types

```{r restype-before, child='Embeds/visual_housing_before.Rmd',echo=FALSE}
```

## Ownership Types

```{r owner-before, child='Embeds/visual_homeowner_before.Rmd', echo=FALSE}
```

```{r multi-before, child='Embeds/visual_multifamily_before.Rmd', echo=FALSE}
```

## Rental Units

```{r units-before, child='Embeds/visual_units_before.Rmd', echo=FALSE}
```

## Average Square Footage

```{r sqfeet-before, child='Embeds/visual_sqftg_before.Rmd', echo=FALSE}
```

# Altadena After

The Eaton Fire burned through the Altadena community for nearly 24 days. Prior research documents how West Altadena was disproportionately devastated by the fire with all fire-related deaths occurring West of Lake Avenue.^[https://www.nbcnews.com/news/us-news/eaton-fire-deaths-los-angeles-evacuation-orders-took-hours-rcna188729] Due to systemic failures in alert and evacuation operations, West Altadena did not receive evacuation orders until nine hours after the Eaton Fire began. A change in wind speed and direction aggressively pushed the fire west early in the morning of January 8th.^[https://file.lacounty.gov/SDSInter/bos/supdocs/207915.pdf] 

We analyzed the destruction from the fire West and East of Lake Avenue examining the share of housing types significantly damaged by the fire. These trends help inform the housing stock lost to the fire and in need of rebuilding in each community; and therefore, also the Altadenans most at risk of displacement.

## Map

West Altadena was largely destroyed by the fire with the fire perimeter covering a significant area of the community. There are pockets of housing that sustained no damage, such as portions along Glenrose Avenue and blocks around Lake Avenue and Dolores Drive. East Altadena had more pockets with no damage such as along North Hollister Avenue, East Mendocino Street, and East Calaveras Street. However, throughout Altadena, even pockets with no damage are surrounded by properties significantly damaged or destroyed. 

```{r map-after, child='Embeds/map_after_eaton.Rmd', echo=FALSE}
```


## Residential Types Lost

```{r restype-after, child='Embeds/visual_housing_after.Rmd', echo=FALSE}
```

```{r multi-after, child='Embeds/visual_multifamily_after.Rmd', echo=FALSE}
```

## Ownership Types Lost

```{r owner-after, child='Embeds/visual_homeowner_after.Rmd', echo=FALSE}
```

## Rental Units Lost

```{r units-after, child='Embeds/visual_units_after.Rmd', echo=FALSE}
```

## Average Square Footage Lost

```{r sqftg-after, child='Embeds/visual_sqftg_after.Rmd', echo=FALSE}
```

# Altadena Rebuilding

The dust from the fire has settled, but Altadena faces a long recovery ahead. To rebuild, varying types of property owners need support in navigating insurance, permits, hazards, and new building standards. Meanwhile, predatory investors have already begun displacing residents and taken advantage of owners when they were most vulnerable. We analyzed the properties that have sold and those that are in progress toward rebuilding as well as the hazard and zoning considerations owners must take into account when rebuilding. All of these factors will affect the path and result of Altadena recovering. Based on September 2025 data from the L.A. County Assessor's office, we mapped and analyzed the properties that have sold. Based on L.A. County permit data we scraped in October 2025, we analyzed the share of significantly damaged properties that are on the road to rebuilding.

## Map

Few properties have sold since the fire, but most are in West Altadena with several homes selling around Lake Avenue and Wapello Street. Meanwhile, most properties have completed debris removal but have not sought an initial permit to rebuild. As properties rebuild, they must also contend with fire and lead hazards in the community. The northern portions of West and East Altadena are in fire hazard zones, with the northern most portions in very high hazard areas that require additional defense measures. Meanwhile, the southern and most western portions of Altadena face higher lead levels--downwind from the fire. Beyond the direct fire perimeter, homes outside of the Eaton Fire burn area are in grids with average lead levels above recommended screening thresholds.

```{r planning-map, child='Embeds/map_damage.Rmd', echo=FALSE}
```

## Properties Sold

```{r sold-month, child='Embeds/visual_sales_bymonth.Rmd', echo=FALSE}
```

```{r sold-damage, child='Embeds/visual_sales_damage.Rmd', echo=FALSE}
```

```{r sold-owner, child='Embeds/visual_sales_owner_renter.Rmd', echo=FALSE}
```

## Permits

```{r permits-area, child='Embeds/visual_permits_damage.Rmd', echo=FALSE}
```

```{r permits-owner, child='Embeds/visual_permits_owner.Rmd', echo=FALSE}
```

## Lead Contamination 

```{r res-lead, child='Embeds/visual_restype_highlead.Rmd', echo=FALSE}
```

```{r owner-lead, child='Embeds/visual_homeowner_highlead.Rmd', echo=FALSE}
```

## Fire Hazards

```{r fhsz, child='Embeds/visual_damage_fhsz.Rmd', echo=FALSE}
```

## Zoning

```{r zoning, child='Embeds/visual_zoning.Rmd', echo=FALSE}
```

# Recommendations (in progress)

* Implement a suite of strategies to support different types of properties and owners in rebuilding. Ensure that taken together policies and programs designed to support Altadena recovery consider what single-family and multifamily homes need as well as what homeowners and landlords from the community need.

* Help safeguard multifamily properties and rental units to prevent displacement of Altadenan renters, particularly with an emphasis on recovering multifamily and rental units in West Altadena that were disproportionately affected.

* Protect homeowners from real estate speculation to preserve homeownership in Altadena. This includes continuing and enforcing bans on unsolicited offers, as enacted by [AB 851](https://calmatters.digitaldemocracy.org/bills/ca_202520260ab851). For property owners who decide they must sell their properties, invest in community or nonprofit acquisition programs, like community land trusts, and prioritize sales to community buyers.^[https://www.saje.net/wp-content/uploads/2025/10/ConfrontingDisaster.10.10.2025.pdf] Together these policies can help prevent speculation, keep properties in the community, and preserve community ownership.

* Track property sales and values in Altadena to understand the long-term impact on future home values and wealth, and respond with appropriate policies to protect against surges or drops in property values that are out-of-sync with L.A. County housing market. 

* Implement policies, practices, and funding to support homeowners seeking to retain and rebuild their properties, such as assistance acquiring permits, finding contractors, and funding construction that exceeds funds available through their insurance policies.

* Track SB 9 permits to ensure Altadena's single-family homes are preserved and not overwhelmingly changed into higher-density multifamily units. At the same time, track progress toward building multifamily units to preserve the relatively few multifamily housing options in Altadena.

* Identify and address hurdles to rebuilding at various stages in permitting process, such as initially compiling necessary plans and paperwork to obtain permits, and then the financing needed to begin and complete construction.

* Provide additional environmental testing and mitigation support to residential properties in high lead testing grids particularly rental units and multifamily properties in East Altadena, and single-family homes in West Altadena. This includes both intact and significantly damaged homes with high lead levels.

* Pursue programs and funding to support Altadenan community members in safely rebuilding their properties in compliance with the state's new defensible space requirements, such as helping to navigate state requirements and finance changes to their properties to address requirements. 

* Pursue broader disaster preparedness for communities within and outside of Fire Hazard Zones given that the Altadena community was mostly outside of the Very High Fire Zone, yet endured one of the worst human- and climate- caused catastrophes in California in the last decade. This can include providing resources for low-income communities and communities of color to prepare their real and personal property for disaster. 

* Address systemic issues in emergency management systems that led to the disorganized response in evacuating Altadena and responding to the shifting fire. This can include investing more general fund money into more sound and robust alerting systems and comparable staffing in the Office of Emergency Management (OEM) as well as updating outdated policies, procedures, and technology to ensure evacuation procedures and emergency personnel work efficiently.^[https://file.lacounty.gov/SDSInter/bos/supdocs/207915.pdf]


# Conclusion (in progress)