---
title: 'Altadena Boundaries and Destroyed Properties'
subtitle: "Operationalizing Boundaries and Damage Categories"
author: "Catalyst California"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 2
    toc_float: yes
---
<!-- link to Google Font -->
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE
)

# Library and environment set up ----

library(sf)
library(dplyr)
# library(highcharter)
library(rmapshaper)
library(leaflet)
library(htmlwidgets)
library(tigris)
# library(tidycensus)
library(stringr)
library(showtext)
library(extrafont)
library(readxl)
library(tidyverse)
library(janitor)
library(flextable)
library(kableExtra)
# library(ggplot2)
# options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))
options(scipen=999)

# Colors and Fonts ----
## COLORS ##
papaya <- "#F25922"
peridot <- "#CEEA01"
lavender <- "#CEC4E2"
meteorite <- "#3A207D"
ccblue <- "#0860BC"

black <- "#000000"
alabaster<-"#FBFBFB"
gainsboro <- "#DEDEDE"

## FONTS ## 

font_add_google("Inter")

# font_import()
loadfonts(device = "win")
windowsFonts()
showtext_auto()

# Load in and prep shapes ----
source("W:\\RDA Team\\R\\credentials_source.R")

con_alt <- connect_to_db("altadena_recovery_rebuild")
con_rda <- connect_to_db("rda_shared_data")
con_fires <- connect_to_db("la_fires")

# eaton fire shape -
eaton_fire <- st_read(con_alt, query="SELECT * FROM data.eaton_fire_prmtr_3310", geom="geom")


# damaged structures database -
eaton_damage <- st_read(con_alt, query="SELECT * FROM data.eaton_fire_dmg_insp_3310", geom="geom")

## filter for only damaged parcels and simplify
damaged_parcels_simple <- eaton_damage %>%
  transform(4326)


# tract xwalk from la fires project -
tracts <- st_read(con_fires, query="SELECT * FROM data.tract_firezone_xwalk where fire_geoid='Eaton'")


# get detailed tract boundaries from tigris to match provided maps -
lac_tracts <- tracts("CA",county="Los Angeles",year=2023) 

## join to lafire xwalk to get prc intersect and detailed tract shapes in one dataframe for altadena
altadena_tracts <- lac_tracts %>% filter(GEOID %in% tracts$tract_geoid) %>% select(GEOID,NAME) %>%
  left_join(tracts %>% select(tract_geoid, prc_tract_area, prc_fire_area), by=c("GEOID"="tract_geoid")) %>%
  # limit to tracts over 5% in burn radius
  filter(prc_tract_area>=.05)

# add an indicator for west/east of lake avenue  and clean up table
names(altadena_tracts ) <- tolower(names(altadena_tracts ))

altadena_tracts <- altadena_tracts %>%
  mutate(location=ifelse(geoid %in% c("06037460301","06037460200","06037461100","06037461000","06037460302","06037461300"), 
                         "West of Lake Ave", "East of Lake Ave"))%>%
  select(geoid,name,location,everything())%>%
  mutate(prc_tract_area=100*as.numeric(prc_tract_area),
         prc_fire_area=100*prc_fire_area) 


# get city boundaries from tigris for mapping -
lac_places <- places(state="CA", year=2023)

## reduce to places around altadena which requires a spatial join with census tracts
## check geometries
st_crs(altadena_tracts)
st_crs(lac_places)
## 4269 which is a coordinate reference system but not a projection, we need to transform to a projection

altadena_tracts <- altadena_tracts %>% st_transform(3310)
lac_places <- lac_places %>% st_transform(3310)

st_crs(lac_places)
## now its a projection

## subset places to those in/around tracts
lac_places_subset <- lac_places[altadena_tracts, op=st_intersects]

## clean up column names
names(lac_places_subset) <- tolower(names(lac_places_subset))


# Make sure all polygons are in 4326
eaton_fire <- eaton_fire %>%
    st_transform(4326)

lac_places_subset <- lac_places_subset %>%
    st_transform(4326)

altadena_tracts <- altadena_tracts %>%
    st_transform(4326)




```

The following explores how to define Altadena boundaries and damage categories for this mapping project. We explore CalFire's Damage Inspection Database to evaluate structure and damage combinations for the parcels that were assessed for damage. We also map the damage data, census tracts, cities, and the Eaton Fire burn area to evaluate the best boundaries to use East and West of Lake Avenue.


# Structure Categories

CalFire's database identifies the category of the structure assessed, e.g., commercial, infrastructure, residential. This table shows the different categories of structures assessed and their damage levels.

```{r damaged-properties}

# explore damaged property characteristics

# need a factor level first
damaged_parcels_simple$damage_ordered <- factor(
      damaged_parcels_simple$damage,
      levels = c("Inaccessible","No Damage", "Affected (1-9%)", "Minor (10-25%)", "Major (26-50%)", "Destroyed (>50%)"), # Define the order of levels
      labels = c("Inaccessible","No Damage", "Affected","Minor","Major","Destroyed")
    )

damaged_parcels_simple %>% st_drop_geometry() %>%
  tabyl(structurec,damage_ordered) %>%
  adorn_totals(c("row","col")) %>%
  adorn_percentages("row")   %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns() %>% 
  kable()

# explore structuret and c combined
structure_type_cat <- table(damaged_parcels_simple$structuret,  damaged_parcels_simple$structurec) %>% as.data.frame()
# structuret for most part nests within structure c

# let's check if any apns have multiple structure categories other than other minor structure
unique_parcel_count <- length(unique(damaged_parcels_simple$apn))
# 11068, use this to make sure we don't go over this number

unique_parcels_addresses <- damaged_parcels_simple %>%
  st_drop_geometry() %>%
  group_by(apn, streetnumb, streetname, streettype, city) %>%
  summarise(structure_count=n(),
            structure_category_list = list(unique(structurec)),
            structure_type_list = list(unique(structuret)),
            damage_types_list=list(unique(damage))) %>%
  ungroup()
# grouping by addresses and apn resulted still in duplicate records, why?
# there are 11520 records here
# let's check where there are duplicates

dup_parcels <- unique_parcels_addresses %>%
  group_by(apn) %>%
  summarise(address_count=n()) %>%
  filter(address_count>=2) %>%
  ungroup () %>%
  left_join(unique_parcels_addresses)
# ones with the most duplicates are condos or larger multifamily properties
# But even ones with some structures destroyed and others not, have misfortunate/calamity status see 5751021036
# 2601 DOVE CREEK LN, NO A still has misfortune/calamity status approved https://portal.assessor.lacounty.gov/parceldetail/5751021004

# let's map single family residences with different damage levels
# let's map multifamily residences with different damage levels
# let's map apns with mixed used separately -- do we include these?

```

## Properties with Mixed Structure Categories
CalFire damage data is at the structure or building level, meaning that each separate structure on a property is given a different structure category and damage assessment. To analyze the data, we will need to identify how best to categorize mixed properties.

Some properties may have mixed structure types, e.g., both a single residence and commercial use. This table shows the combination categories for unique parcel numbers with multiple structure categories. Assessor data will help us filter for parcels currently being used for residential in cases like these where multiple categories are recorded. Most cases with two or more category types are due to a minor structure and a single or multiple residence. Other minor structures tend to be garages from our mapping checks online. Rather than developing a typology with CalFire's data, we will select residential properties based on use codes from the L.A. County Assessor's office. 

```{r}
# go back to structure types, how can we filter for structures that are residential? What are the combinations?
# create a single address field
damaged_parcels_simple <- damaged_parcels_simple %>%
  mutate(full_address=paste0(streetnumb, " ", streetname, " ", streettype, ", ", city))

# get a list of parcels with different damage levels and structure categories
unique_parcels <- damaged_parcels_simple %>%
  st_drop_geometry() %>%
  group_by(apn) %>%
  summarise(structure_count=n(),
            structure_category_list = list(unique(structurec)),
            structure_category_count=length(unique(structure_category_list[[1]])),
            structure_type_list = list(unique(structuret)),
            structure_type_count=length(unique(structure_type_list[[1]])),
            damage_type_list=list(unique(damage)),
            damage_type_count=length(unique(damage_type_list[[1]])),
            address_list = list(unique(full_address)),
            address_count=length(unique(address_list[[1]]))) %>%
  ungroup()

# sort list of structure category for easier grouping later
unique_parcels$structure_category_list <- lapply(unique_parcels$structure_category_list,sort)

# sort list of damage for easier grouping later
unique_parcels$damage_type_list <- lapply(unique_parcels$damage_type_list,sort)

unique_parcels %>%
  st_drop_geometry() %>%
  filter(structure_category_count>=2) %>%
group_by(structure_category_list) %>%
  summarize(count=n(),
            structure_category_label=toString(unlist(unique(structure_category_list)))) %>%
  ungroup() %>%
  select(structure_category_label, count) %>%
  arrange(count) %>%
  kable()

# approximately less than 50 have some sort of mix with residential/nonresidential/infrastructure


```



# Properties with Mixed Damage Levels
Each structure on a property is given a separate damage level. One property could have a structure marked as 'Destroyed' and another structure marked as 'No Damage'.
How do we treat these mixed damage level properties?

This first table shows the combination of damage levels for properties with at least one residential use and two or more damage levels. The majoried of mixed damage levels include Destroyed + Affected or Destroyed + No Damage. The next most common are Affected + No Damage.
```{r map-single-residential-properties-multiple-damage}

# what's the target list to include
target <- list("Mixed Commercial/Residential","Multiple Residence","Single Residence")

# filter for parcels that have more than 2 damage levels and that have at least one identified residential use
df_any_res <- unique_parcels %>%
  filter(damage_type_count>=2) %>%
  mutate(target_check = map_lgl(structure_category_list, ~any((. %in% target)))) %>%
  # filter for only parcels where at least one category is in our list
  filter(target_check == TRUE)


# tabulate the combo damage levels
df_any_res %>%
  st_drop_geometry() %>%
  group_by(damage_type_list) %>%
  summarize(count=n(),
            damage_type_label=toString(unlist(unique(damage_type_list)))) %>%
  ungroup() %>%
  select(damage_type_label, count) %>%
  arrange(count) %>%
  kable()

# take out a couple of records that are inaccessible & no damage
unique_parcels <- unique_parcels %>%
  filter(!as.character(damage_type_list) == 'c("Inaccessible", "No Damage")')
# 11066 records left

```


## Single Residences with Multiple Damage Levels
This map plots the properties labeled as single residences that have multiple damage levels. This is mapped to evaluate whether these properties with mixed damage levels are liveable or not. How should they be treated in the analysis?
```{r}

# what's the target list for this map -- have to include other minor structure due to detached garages
target <- list("Single Residence", "Other Minor Structure")

# filter for parcels that have more than 2 damage levels and that are single residences
df_single <- unique_parcels %>%
  filter(damage_type_count>=2) %>%
  mutate(target_check = map_lgl(structure_category_list, ~any(!(. %in% target)))) %>%
  # filter for only parcels where category is in our target list--excludes any case where at least one category doesn't meet our list
  filter(target_check == FALSE)

# there are a couple of apns with just other minor structure but they appear to be single family from the parcel maps: see 5751001003
# https://portal.assessor.lacounty.gov/parceldetail/5751001003

# tabulate the combo damage levels
df_single %>%
  st_drop_geometry() %>%
group_by(damage_type_list) %>%
  summarize(count=n(),
            damage_type_label=toString(unlist(unique(damage_type_list)))) %>%
  ungroup() %>%
  select(damage_type_label, count) %>%
  arrange(count)
# Mostly 
# Affected (1-9%), No Damage	182			
# Affected (1-9%), Destroyed (>50%)	216			
# Destroyed (>50%), No Damage	387	

# revise factor level for this map and others from now on -- using original dataframe for mapping that is ungrouped so we can see locations and distinctions
damaged_parcels_simple$damage_ordered <- factor(
      damaged_parcels_simple$damage,
      levels = c("Inaccessible","No Damage", "Affected (1-9%)", "Minor (10-25%)", "Major (26-50%)", "Destroyed (>50%)"), # Define the order of levels
      labels = c("Inaccessible","No Damage", "Affected","Minor","Major","Destroyed")
    )

pal <- colorFactor(palette = c("#6F6F6F","#909078","#177FEB", "#4886BD", "#CC983F","#FF9E0D"), domain = damaged_parcels_simple$damage_ordered)


# filter for these properties single resident properties
df_map <- damaged_parcels_simple %>%
  filter(apn %in% df_single$apn)


leaflet () %>%
addProviderTiles(providers$CartoDB.Positron) %>%
    setView(-118.104631, 34.185104, zoom = 12) %>%
  addPolygons(data=eaton_fire,
              fillColor="white",
              fillOpacity=0,
              color=meteorite,
              weight=2.5,
              opacity=1,
              popup=~fire_name,
              group="Eaton Fire Perimeter") %>%
      addPolygons(data=lac_places_subset,
              fillColor="white",
              fillOpacity=0,
              color=peridot,
              weight=1.5,
              opacity=1,
              popup=~name,
              group="Cities") %>%
      addCircleMarkers(data=df_map,
        lng = ~longitude,
        lat = ~latitude,
        fillColor = ~pal(damage_ordered), # Color based on category
        radius = 2,
        stroke=TRUE,
        # fill=TRUE,
        fillOpacity = 1,
        color=gainsboro,
        weight=.5,
        opacity=.5,
        popup = ~paste0(damage,"</br>",
                        "Parcel #: ", apn, "</br>",
                        "Structure Category: ", structurec, "</br>",
                        "Address: ", full_address, "</br>",
                        "City: ", city),
        group="Single Residences with Multiple Damage Levels"
      ) %>%
      addLegend(
        position = "bottomright",
        pal = pal,
        values = df_map$damage_ordered,
        title = "Damage level"
      ) %>%
  addLayersControl(
          overlayGroups = c("Eaton Fire Perimeter", "Single Residences with Multiple Damage Levels","Cities"),
        options = layersControlOptions(collapsed = FALSE)
  )


```

I randomly selected 5 properties from the list to explore their damage combos. Most even with only affected damage still have misfortune and calamity status approved for their tax bills.
```{r}
# randomly select some parcels to test
sample <- sample_n(df_single,5)
# retrieved parcels below

```

Parcel Number: 5835017027

Notes: A standing house with destroyed structures behind it and calamity status approved.

https://portal.assessor.lacounty.gov/parceldetail/5835017027

Parcel Number: 5840022039

Notes: Looks completely destroyed though it's only marked as Affected and No Damage. Calamity status is approved.

https://portal.assessor.lacounty.gov/parceldetail/5840022039

Parcel Number: 5857014006

Notes: Unclear what structured on the property has been affected. Calamity status approved. Marked as Affected and No Damage.

https://portal.assessor.lacounty.gov/parceldetail/5857014006

Parcel Number: 5845031013

Notes: Garage is destroyed. Unclear how house is damaged. Calamity status is approved.

https://portal.assessor.lacounty.gov/parceldetail/5845031013

Parcel Number: 5828008010

Notes: Minor damage noted, but garage and house seem fully intact. House is a block from Lincoln Avenue. Calamity status approved.

https://portal.assessor.lacounty.gov/parceldetail/5828008010

## Multiple Residences with Multiple Damage Levels
This map plots the properties labeled as multiple residences that have multiple damage levels. A few that have no or affected damage are noteably outside of the fire perimeter.
```{r}

# what's the target list for this map -- have to include other minor structure due to detached garages
target <- list("Multiple Residence", "Other Minor Structure")

# filter for parcels that have more than 2 damage levels and that are single residences
df_multiple <- unique_parcels %>%
  filter(damage_type_count>=2) %>%
  mutate(target_check = map_lgl(structure_category_list, ~any(!(. %in% target)))) %>%
  # filter for only parcels where category is in our target list--excludes any case where at least one category doesn't meet our list
  filter(target_check == FALSE)

# take out those only labels as other minor structure--they would have been mapped in previous one
# take out a couple of records that are inaccessible & no damage
df_multiple <- df_multiple %>%
  filter(!as.character(structure_category_list) == 'Other Minor Structure')


# # tabulate the combo damage levels
# df_multiple %>%
#   st_drop_geometry() %>%
# group_by(damage_type_list) %>%
#   summarize(count=n(),
#             damage_type_label=toString(unlist(unique(damage_type_list)))) %>%
#   ungroup() %>%
#   select(damage_type_label, count) %>%
#   arrange(count)
# Mostly 
# Destroyed (>50%), Major (26-50%)	3			
# Affected (1-9%), Destroyed (>50%), No Damage	3			
# Destroyed (>50%), No Damage	6			
# Affected (1-9%), No Damage	6	

# filter for these properties multiple resident properties
df_map <- damaged_parcels_simple %>%
  filter(apn %in% df_multiple$apn)


leaflet () %>%
addProviderTiles(providers$CartoDB.Positron) %>%
    setView(-118.104631, 34.185104, zoom = 12) %>%
  addPolygons(data=eaton_fire,
              fillColor="white",
              fillOpacity=0,
              color=meteorite,
              weight=2.5,
              opacity=1,
              popup=~fire_name,
              group="Eaton Fire Perimeter") %>%
      addPolygons(data=lac_places_subset,
              fillColor="white",
              fillOpacity=0,
              color=peridot,
              weight=1.5,
              opacity=1,
              popup=~name,
              group="Cities") %>%
      addCircleMarkers(data=df_map,
        lng = ~longitude,
        lat = ~latitude,
        fillColor = ~pal(damage_ordered), # Color based on category
        radius = 2,
        stroke=TRUE,
        # fill=TRUE,
        fillOpacity = 1,
        color=gainsboro,
        weight=.5,
        opacity=.5,
        popup = ~paste0(damage,"</br>",
                        "Parcel #: ", apn, "</br>",
                        "Structure Category: ", structurec, "</br>",
                        "Address: ", full_address, "</br>",
                        "City: ", city),
        group="Multiple Residences with Multiple Damage Levels"
      ) %>%
      addLegend(
        position = "bottomright",
        pal = pal,
        values = df_map$damage_ordered,
        title = "Damage level"
      ) %>%
  addLayersControl(
          overlayGroups = c("Eaton Fire Perimeter", "Multiple Residences with Multiple Damage Levels","Cities"),
        options = layersControlOptions(collapsed = FALSE)
  )


```
I randomly selected 5 properties from the list to explore their damage combination. Most even with only affected damage still have misfortune and calamity status approved for their tax bills.
```{r}
# randomly select some parcels to test
sample <- sample_n(df_multiple,5)
# retrieved parcels below
```

Parcel Number: 5840015030

Notes: Condo with a couple of destroyed buildings on it and calamity status approved. All 5 structures share the same parcel number.

https://portal.assessor.lacounty.gov/parceldetail/5840015030

Parcel Number: 5853008008

Notes: Mulit-family property with three units. Property is outside of burn area. Calamity status is approved, but the units seem to be intact. They are noted as sustaining no or affected damage.

https://portal.assessor.lacounty.gov/parceldetail/5853008008

Parcel Number: 5845016006

Notes: Multi-family property with three units. Calamity status is approved and the property is completely gone.

https://portal.assessor.lacounty.gov/parceldetail/5845016006

Parcel Number: 5751008029

Notes: Condo complex with one building destroyed. Calamity status is approved, and property is delinquent on taxes. Complex is all on one parcel.

https://portal.assessor.lacounty.gov/parceldetail/5751008029

Parcel Number: 5844001010

Notes: Multi-family property with one building destroyed and calamity status approved.

https://portal.assessor.lacounty.gov/parceldetail/5844001010 


# Proposed Damage Categories
To focus on properties that sustained greater damage and make the data findings more easily interpretable (by including fewer categories), we can group damage levels as follows:

* No Damage, Inaccessible, Affected alone or in combination in each other would be considered as Little to No Damage.
* Minor Damage alone or in combination with No Damage, Inaccessible, or Affected would be considered Some Damage.
* Major Damage or Destroyed alone or in combination with another category would be considered Significant Damage or Destroyed.

## Little to No Damage
These properties with at least some residential use would be considered as little to no damage. They sustained no or affected damage.
```{r}

# what's the target list 
target <- list("No Damage", "Inaccessible", "Affected (1-9%)")

# filter for parcels that only have these damage levels
df_no_damage <- df_any_res %>%
  mutate(target_check = map_lgl(damage_type_list, ~any(!(. %in% target)))) %>%
  # filter for only parcels where no other damage type turns up
  filter(target_check == FALSE)

# filter for these properties to mape
df_map <- damaged_parcels_simple %>%
  filter(apn %in% df_no_damage$apn)


leaflet () %>%
addProviderTiles(providers$CartoDB.Positron) %>%
    setView(-118.104631, 34.185104, zoom = 12) %>%
  addPolygons(data=eaton_fire,
              fillColor="white",
              fillOpacity=0,
              color=meteorite,
              weight=2.5,
              opacity=1,
              popup=~fire_name,
              group="Eaton Fire Perimeter") %>%
      addPolygons(data=lac_places_subset,
              fillColor="white",
              fillOpacity=0,
              color=peridot,
              weight=1.5,
              opacity=1,
              popup=~name,
              group="Cities") %>%
      addCircleMarkers(data=df_map,
        lng = ~longitude,
        lat = ~latitude,
        fillColor = ~pal(damage_ordered), # Color based on category
        radius = 2,
        stroke=TRUE,
        # fill=TRUE,
        fillOpacity = 1,
        color=gainsboro,
        weight=.5,
        opacity=.5,
        popup = ~paste0(damage,"</br>",
                        "Parcel #: ", apn, "</br>",
                        "Structure Category: ", structurec, "</br>",
                        "Address: ", full_address, "</br>",
                        "City: ", city),
        group="Residences with Little to No Damage"
      ) %>%
      addLegend(
        position = "bottomright",
        pal = pal,
        values = df_map$damage_ordered,
        title = "Damage level"
      ) %>%
  addLayersControl(
          overlayGroups = c("Eaton Fire Perimeter", "Residences with Little to No Damage","Cities"),
        options = layersControlOptions(collapsed = FALSE)
  )

```


## Some Damage
These properties with at least some residential use would be considered as sustaining some damage. They sustained at most minor damage (10-25%) and more than affected damage (1-9%).
```{r}

# what's the target list 
target <- list("No Damage", "Inaccessible", "Affected (1-9%)","Minor (10-25%)")

# filter for parcels that only have these damage levels
df_some_damage <- df_any_res %>%
  mutate(target_check = map_lgl(damage_type_list, ~any(!(. %in% target)))) %>%
  # filter for only parcels where no other damage type turns up
  filter(target_check == FALSE)

# filter for these properties to map and exclude the ones from the first 
df_map <- damaged_parcels_simple %>%
  filter(apn %in% df_some_damage$apn) %>%
  filter(!apn %in% df_no_damage$apn) 


leaflet () %>%
addProviderTiles(providers$CartoDB.Positron) %>%
    setView(-118.104631, 34.185104, zoom = 12) %>%
  addPolygons(data=eaton_fire,
              fillColor="white",
              fillOpacity=0,
              color=meteorite,
              weight=2.5,
              opacity=1,
              popup=~fire_name,
              group="Eaton Fire Perimeter") %>%
      addPolygons(data=lac_places_subset,
              fillColor="white",
              fillOpacity=0,
              color=peridot,
              weight=1.5,
              opacity=1,
              popup=~name,
              group="Cities") %>%
      addCircleMarkers(data=df_map,
        lng = ~longitude,
        lat = ~latitude,
        fillColor = ~pal(damage_ordered), # Color based on category
        radius = 2,
        stroke=TRUE,
        # fill=TRUE,
        fillOpacity = 1,
        color=gainsboro,
        weight=.5,
        opacity=.5,
        popup = ~paste0(damage,"</br>",
                        "Parcel #: ", apn, "</br>",
                        "Structure Category: ", structurec, "</br>",
                        "Address: ", full_address, "</br>",
                        "City: ", city),
        group="Residences with Some Damage"
      ) %>%
      addLegend(
        position = "bottomright",
        pal = pal,
        values = df_map$damage_ordered,
        title = "Damage level"
      ) %>%
  addLayersControl(
          overlayGroups = c("Eaton Fire Perimeter", "Residences with Some Damage","Cities"),
        options = layersControlOptions(collapsed = FALSE)
  )

```

## Significant Damage or Destroyed
These properties with at least some residential use would be considered as sustaining significant damage or were destroyed. They sustained at least major damage (26-50%).
```{r}

# what's the target list 
target <- list("Destroyed (>50%)", "Major (26-50%)")

# filter for parcels that only have these damage levels
df_destroyed <- df_any_res %>%
  mutate(target_check = map_lgl(damage_type_list, ~any((. %in% target)))) %>%
  # filter for only parcels where at least one of these damage levels turns up
  filter(target_check == TRUE)

# filter for these properties to map and exclude the ones from the first 
df_map <- damaged_parcels_simple %>%
  filter(apn %in% df_destroyed$apn)


leaflet () %>%
addProviderTiles(providers$CartoDB.Positron) %>%
    setView(-118.104631, 34.185104, zoom = 12) %>%
  addPolygons(data=eaton_fire,
              fillColor="white",
              fillOpacity=0,
              color=meteorite,
              weight=2.5,
              opacity=1,
              popup=~fire_name,
              group="Eaton Fire Perimeter") %>%
      addPolygons(data=lac_places_subset,
              fillColor="white",
              fillOpacity=0,
              color=peridot,
              weight=1.5,
              opacity=1,
              popup=~name,
              group="Cities") %>%
      addCircleMarkers(data=df_map,
        lng = ~longitude,
        lat = ~latitude,
        fillColor = ~pal(damage_ordered), # Color based on category
        radius = 2,
        stroke=TRUE,
        # fill=TRUE,
        fillOpacity = 1,
        color=gainsboro,
        weight=.5,
        opacity=.5,
        popup = ~paste0(damage,"</br>",
                        "Parcel #: ", apn, "</br>",
                        "Structure Category: ", structurec, "</br>",
                        "Address: ", full_address, "</br>",
                        "City: ", city),
        group="Residences with Some Damage"
      ) %>%
      addLegend(
        position = "bottomright",
        pal = pal,
        values = df_map$damage_ordered,
        title = "Damage level"
      ) %>%
  addLayersControl(
          overlayGroups = c("Eaton Fire Perimeter", "Residences with Some Damage","Cities"),
        options = layersControlOptions(collapsed = FALSE)
  )

```

# Eaton Fire Boundary Compared to Damaged/Destroyed Properties, Cities, and Tract Boundaries
Map only displays residential or mixed use properties that sustained at least an "affected (1-9%)" damage level. Kinneloa Mesa did not sustain any damage. Notably, some properties marked as destroyed fall just slightly outside of the burn area, meaning when selecting parcels from the Assessor data to analyze, we might need to rely on parcel numbers or addresses to select parcels rather than joins to the burn area. The burn area might intersect with the parcel boundaries. Only parcels in the burn area or sustaining at least some damage would be included in the detailed analysis.

```{r altadena-map}

# map damaged parcels and check boundaries we want to use against tracts and places

# for purpose of boundary mapping, let's just map properties that were damaged, regardless if they are duplicates and let's focus on residential propertieis
only_damaged_parcels <- damaged_parcels_simple %>%
  filter(!damage %in% c("No Damage", "Inaccessible")) %>%
  filter(structurec %in% c("Mixed Commercial/Residential","Multiple Residence","Single Residence","Other Minor Structure"))

# revise factor level
only_damaged_parcels$damage_ordered <- factor(
      only_damaged_parcels$damage,
      levels = c("Affected (1-9%)", "Minor (10-25%)", "Major (26-50%)", "Destroyed (>50%)"), # Define the order of levels
      labels = c("Affected","Minor","Major","Destroyed")
    )

# Create a color palette based on the damage level
# table(only_damaged_parcels$damage,useNA='always') # get values again to assign colors

pal <- colorFactor(palette = c("#177FEB", "#4886BD", "#CC983F","#FF9E0D"), domain = only_damaged_parcels$damage_ordered)
# table(only_damaged_parcels$damage_ordered,useNA='always')

leaflet () %>%
addProviderTiles(providers$CartoDB.Positron) %>%
    setView(-118.104631, 34.185104, zoom = 12) %>%
  addPolygons(data=eaton_fire,
              fillColor="white",
              fillOpacity=0,
              color=meteorite,
              weight=2.5,
              opacity=1,
              popup=~"Eaton Fire Burn Area",
              group="Eaton Fire Perimeter") %>%
    addPolygons(data=altadena_tracts,
              fillColor="white",
              fillOpacity=0,
              color=black,
              weight=1,
              opacity=1,
              popup=~name,
              group="Census Tracts") %>%
      addPolygons(data=lac_places_subset,
              fillColor="white",
              fillOpacity=0,
              color=peridot,
              weight=1.5,
              opacity=1,
              popup=~name,
              group="Cities") %>%
      addCircleMarkers(data=only_damaged_parcels,
        lng = ~longitude,
        lat = ~latitude,
        fillColor = ~pal(damage_ordered), # Color based on category
        radius = 2,
        stroke=TRUE,
        # fill=TRUE,
        fillOpacity = 1,
        color=gainsboro,
        weight=.5,
        opacity=.5,
        popup = ~paste0(damage,"</br>",
                        "Parcel #: ", apn, "</br>",
                        "Structure Category: ", structurec, "</br>",
                        "Address: ", full_address, "</br>",
                        "City: ", city),
                        # Optional: Add a popup
        group="Damaged Properties"
      ) %>%
      addLegend(
        position = "bottomright",
        pal = pal,
        values = only_damaged_parcels$damage_ordered,
        title = "Damage level"
      ) %>%
  addLayersControl(
          overlayGroups = c("Eaton Fire Perimeter", "Census Tracts","Damaged Properties","Cities"),
        options = layersControlOptions(collapsed = FALSE)
  )
      



```