---
title: 'Altadena Boundaries and Destroyed Properties'
subtitle: "Operationalizing Boundaries and Damage Categories"
author: "Catalyst California"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: "W:\\RDA Team\\R\\cc_brandguide.css"
    toc: yes
    toc_depth: 2
    toc_float: yes
---
<!-- link to Google Font -->
<link href="https://fonts.googleapis.com/css2?family=Inter:wght@100;200;300;400;500;600;700;800;900&display=swap" rel="stylesheet">

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = FALSE, message = FALSE, warning = FALSE, echo = FALSE
)

# Library and environment set up ----

library(sf)
library(dplyr)
# library(highcharter)
library(rmapshaper)
library(leaflet)
library(htmlwidgets)
library(tigris)
# library(tidycensus)
library(stringr)
library(showtext)
library(extrafont)
library(readxl)
library(tidyverse)
library(janitor)
library(flextable)
library(kableExtra)
# library(ggplot2)
# options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))
options(scipen=999)

# Colors and Fonts ----
## COLORS ##
papaya <- "#F25922"
peridot <- "#CEEA01"
lavender <- "#CEC4E2"
meteorite <- "#3A207D"
ccblue <- "#0860BC"

black <- "#000000"
alabaster<-"#FBFBFB"
gainsboro <- "#DEDEDE"

## FONTS ## 

font_add_google("Inter")

# font_import()
loadfonts(device = "win")
windowsFonts()
showtext_auto()

# Load in and prep shapes ----
source("W:\\RDA Team\\R\\credentials_source.R")

con_alt <- connect_to_db("altadena_recovery_rebuild")
con_rda <- connect_to_db("rda_shared_data")
con_fires <- connect_to_db("la_fires")

# eaton fire shape -
eaton_fire <- st_read(con_alt, query="SELECT * FROM data.eaton_fire_prmtr_3310", geom="geom")


# damaged structures database -
eaton_damage <- st_read(con_alt, query="SELECT * FROM data.eaton_fire_dmg_insp_3310", geom="geom") %>% rename(structurec=structure_category,                                                              structuret=structure_type,
                                                                                                              apn=apn_parcel)

## filter for only damaged parcels and simplify
damaged_parcels_simple <- eaton_damage %>%
  transform(4326)


# tract xwalk from la fires project -
tracts <- st_read(con_fires, query="SELECT * FROM data.tract_firezone_xwalk where fire_geoid='Eaton'")


# get detailed tract boundaries from tigris to match provided maps -
lac_tracts <- tracts("CA",county="Los Angeles",year=2023) 

## join to lafire xwalk to get prc intersect and detailed tract shapes in one dataframe for altadena
altadena_tracts <- lac_tracts %>% filter(GEOID %in% tracts$tract_geoid) %>% select(GEOID,NAME) %>%
  left_join(tracts %>% select(tract_geoid, prc_tract_area, prc_fire_area), by=c("GEOID"="tract_geoid")) %>%
  # limit to tracts over 5% in burn radius
  filter(prc_tract_area>=.05)

# add an indicator for west/east of lake avenue  and clean up table
names(altadena_tracts ) <- tolower(names(altadena_tracts ))

altadena_tracts <- altadena_tracts %>%
  mutate(location=ifelse(geoid %in% c("06037460301","06037460200","06037461100","06037461000","06037460302","06037461300"), 
                         "West of Lake Ave", "East of Lake Ave"))%>%
  select(geoid,name,location,everything())%>%
  mutate(prc_tract_area=100*as.numeric(prc_tract_area),
         prc_fire_area=100*prc_fire_area) 


# get city boundaries from tigris for mapping -
lac_places <- places(state="CA", year=2023)

## reduce to places around altadena which requires a spatial join with census tracts
## check geometries
st_crs(altadena_tracts)
st_crs(lac_places)
## 4269 which is a coordinate reference system but not a projection, we need to transform to a projection

altadena_tracts <- altadena_tracts %>% st_transform(3310)
lac_places <- lac_places %>% st_transform(3310)

st_crs(lac_places)
## now its a projection

## subset places to those in/around tracts
lac_places_subset <- lac_places[altadena_tracts, op=st_intersects]

## clean up column names
names(lac_places_subset) <- tolower(names(lac_places_subset))


# Make sure all polygons are in 4326
eaton_fire <- eaton_fire %>%
    st_transform(4326)

lac_places_subset <- lac_places_subset %>%
    st_transform(4326)

altadena_tracts <- altadena_tracts %>%
    st_transform(4326)

```

The following explores how to define Altadena boundaries and damage categories. We explore CalFire's Damage Inspection Database to evaluate structure and damage combinations for the parcels in or near the Eaton Fire burn area. We also map the damaged data, census tracts, cities, and burn area to evaluate the best boundaries to use East and West of Lake Avenue for future comparisons about housing stock.


# Structure Categories

CalFire's database identifies the category of the structure assessed, e.g., commercial, infrastructure, residential. This table shows the different categories of assessed structures and their damage levels. This table does not account for unique property parcels. In other words, a property could have a minor structure and a single residence, each with their own category and damage level.

```{r structures}

# explore damaged property characteristics

# need a factor level first
damaged_parcels_simple$damage_ordered <- factor(
      damaged_parcels_simple$damage,
      levels = c("Inaccessible","No Damage", "Affected (1-9%)", "Minor (10-25%)", "Major (26-50%)", "Destroyed (>50%)"), # Define the order of levels
      labels = c("Inaccessible","No Damage", "Affected","Minor","Major","Destroyed")
    )

# look at damage level by structure category
damaged_parcels_simple %>% st_drop_geometry() %>%
  rename(Category=structurec) %>%
  tabyl(Category,damage_ordered) %>%
  adorn_totals(c("row","col")) %>%
  adorn_percentages("row")   %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() %>% 
  kable() %>%
  column_spec(2:7, width='25mm')

# explore structuret and c combined - how do they relate?
structure_type_cat <- table(damaged_parcels_simple$structuret,  damaged_parcels_simple$structurec) %>% as.data.frame()
# structuret for most part nests within structure c

# let's check if any apns have multiple structure categories other than other minor structure
unique_parcel_count <- length(unique(damaged_parcels_simple$apn))
# 11068, use this to make sure we don't go over or under this number when checking for duplicates

# unique parcels by apn and address
unique_parcels_addresses <- damaged_parcels_simple %>%
  st_drop_geometry() %>%
  group_by(apn, street_number, street_name, street_type, city) %>%
  summarise(structure_count=n(),
            structure_category_list = list(unique(structurec)),
            structure_type_list = list(unique(structuret)),
            damage_types_list=list(unique(damage))) %>%
  ungroup()
# grouping by addresses and apn resulted still in duplicate records, why?
# there are 11520 records here
# let's check where there are duplicates

dup_parcels <- unique_parcels_addresses %>%
  group_by(apn) %>%
  summarise(address_count=n()) %>%
  filter(address_count>=2) %>%
  ungroup () %>%
  left_join(unique_parcels_addresses)%>%
  arrange(desc(address_count))

View(dup_parcels)
# ones with the most duplicate addresses on a single parcel are condos or larger multifamily properties
# But even ones with some structures destroyed and others not, have misfortunate/calamity status see 5751021036
# 2601 DOVE CREEK LN, NO A still has misfortune/calamity status approved https://portal.assessor.lacounty.gov/parceldetail/5751021004
# APN coded as single family, but is a condo --
# Coded in CalFire as 5833019042, but this parcel was deleted. An example address on that parcel is 155 E PALM ST - https://portal.assessor.lacounty.gov/parceldetail/5833019043

# like other data, single residences have highest prevalence in duplicates--though some of these are still condos and some may be ADUs where there are 2 addresses recorded in CalFire https://portal.assessor.lacounty.gov/parceldetail/5751012006

# let's explore and map single family residences with different damage levels
# let's explore and map multifamily residences with different damage levels

```

## Properties with Mixed Structure Categories
CalFire damage data is provided at the structure or building level. In other words, each separate structure on a property is given a different structure category and damage assessment. To analyze the data, we will need to identify how best to categorize mixed properties.

Some properties may have mixed structure types, e.g., both a single residence and commercial use. This table shows the combination categories for unique parcels with multiple structures and structure categories. L.A. County Assessor data will help us filter for parcels currently being used for residential or mixed use in cases like these where multiple categories are recorded. 

Most cases with two or more category types are due to a minor structure and a single or multiple residence. Minor structures tend to be garages from our mapping checks online. Rather than developing a typology from CalFire's data, we will select residential properties based on use codes from the L.A. County Assessor's office. We will largely be focusing on properties with residential use codes, but may included mixed-use properties if we are able to distinguish the square footage/units lost particular to the residential part.

```{r structures-mixed}

# first go back to structure types, how can we filter for structures that are residential? What are the combinations? Or do we rely on assessor data?
# create a single address field
damaged_parcels_simple <- damaged_parcels_simple %>%
  mutate(full_address=paste0(street_number, " ", street_name, " ", street_type, ", ", city))

# get a list of parcels with different damage levels and structure categories, keep addresses
unique_parcels <- damaged_parcels_simple %>%
  st_drop_geometry() %>%
  group_by(apn) %>%
  summarise(structure_count=n(),
            structure_category_list = list(unique(structurec)),
            structure_category_count=length(unique(structure_category_list[[1]])),
            structure_type_list = list(unique(structuret)),
            structure_type_count=length(unique(structure_type_list[[1]])),
            damage_type_list=list(unique(damage)),
            damage_type_count=length(unique(damage_type_list[[1]])),
            address_list = list(unique(full_address)),
            address_count=length(unique(address_list[[1]]))) %>%
  ungroup() %>%
  mutate(mixed_damage=ifelse(damage_type_count<2, "One Damage Type",
                             "Two or More"))

# check recoding of mixed damage category, looks good
# table(unique_parcels$damage_type_count,unique_parcels$mixed_damage)

# sort list of structure category for easier grouping later
unique_parcels$structure_category_list <- lapply(unique_parcels$structure_category_list,sort)

# sort list of damage for easier grouping later
unique_parcels$damage_type_list <- lapply(unique_parcels$damage_type_list,sort)

# properties with combined structure category types--what are the combos?
unique_parcels %>%
  st_drop_geometry() %>%
  filter(structure_category_count>=2) %>%
group_by(structure_category_list) %>%
  summarize(count=n(),
            structure_category_label=toString(unlist(unique(structure_category_list)))) %>%
  ungroup() %>%
  select(structure_category_label, count) %>%
  arrange(count) %>%
  rename(Categories=structure_category_label,
         Count=count) %>%
  kable()

# approximately less than 50 have some sort of mix with residential/nonresidential/infrastructure, most are either residential or commercial
# we'll probably just use assessor data to filter for use codes of residential, might look for mixed commercial but we may be unable to distinguish unit size lost for those unless it's easy to distinguish the unit size lost separate from commercial square footage


```



# Damage Categories
Each structure on a property is given a separate damage level. One property could have a structure marked as 'Destroyed' and another structure marked as 'No Damage'. We will need to analyze the properties lost to the fire and in need of rebuilding. To do this, we will need to determine a typology for categorizing a property as lost. L.A. County Assessor data that includes the property square footage and units is provided at the parcel level, which would make matching CalFire's structure-level damage data challenging. Instead, we will use CalFire's damage data to distinguish between residential parcels still standing and not in need of significant rebuilding versus those that are significantly damaged or destroyed and in need of rebuilding.

This first table shows the percentage of properties with combined damage levels by combined structure category. Only 9% of unique parcels in CalFire's database have two or more damage types, but this percentage increases for certain categories. For instance, 16% of single residences with a minor structure on the property have mixed damage levels. 

```{r}
# what percentage of each structure category has mixed damage levels just for any residential use
# what's the target list to include
target <- list("Mixed Commercial/Residential","Multiple Residence","Single Residence")

# filter for parcels that have at least one identified residential use
df_any_res_all <- unique_parcels %>%
  mutate(target_check = map_lgl(structure_category_list, ~any((. %in% target)))) %>%
  # filter for only parcels where at least one category is in our list
  filter(target_check == TRUE)

# look at damage level by structure category
df_any_res_all  %>% st_drop_geometry() %>%
  group_by(structure_category_list) %>%
  mutate(`Structure Category`=toString(unlist(unique(structure_category_list)))) %>%
  tabyl(`Structure Category`, mixed_damage) %>%
  adorn_totals(c("row","col")) %>%
  adorn_percentages("row")   %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() %>%
  kable() 



```

This second table shows the combination of damage levels for properties with at least one residential use and two or more damage levels. The majority of mixed damage levels include Destroyed and Affected or Destroyed and No Damage. The next most common are Affected and No Damage.

```{r damage}

# what's the target list to include
target <- list("Mixed Commercial/Residential","Multiple Residence","Single Residence")

# filter for parcels that have more than 2 damage levels and that have at least one identified residential use
df_any_res_mixed <- unique_parcels %>%
  filter(damage_type_count>=2) %>%
  mutate(target_check = map_lgl(structure_category_list, ~any((. %in% target)))) %>%
  # filter for only parcels where at least one category is in our list
  filter(target_check == TRUE)


# tabulate the combo damage levels for some residential use properties
df_any_res_mixed %>%
  st_drop_geometry() %>%
  group_by(damage_type_list) %>%
  summarize(count=n(),
            damage_type_label=toString(unlist(unique(damage_type_list)))) %>%
  ungroup() %>%
  select(damage_type_label, count) %>%
  rename('Damage Type'=damage_type_label,
         Count=count) %>%
  arrange(Count) %>%
  kable()

# take out a couple of records that are inaccessible & no damage
unique_parcels <- unique_parcels %>%
  filter(!as.character(damage_type_list) == 'c("Inaccessible", "No Damage")')
# 11066 records left

```


## Single Residences with Multiple Damage Levels
This map plots the properties labeled as single residences that have multiple damage levels. This is mapped to evaluate whether these properties with mixed damage levels are liveable or not.

```{r single-mixed}

# what's the target list for this map -- have to include other minor structure due to detached garages
target <- list("Single Residence", "Other Minor Structure")

# filter for parcels that have more than 2 damage levels and that are single residences
df_single <- unique_parcels %>%
  filter(damage_type_count>=2) %>%
  mutate(target_check = map_lgl(structure_category_list, ~any(!(. %in% target)))) %>%
  # filter for only parcels where category is in our target list--excludes any case where at least one category doesn't meet our list
  filter(target_check == FALSE)

# there are a couple of apns with just other minor structure but they appear to be single family from the parcel maps: see 5751001003
# https://portal.assessor.lacounty.gov/parceldetail/5751001003

# # tabulate the combo damage levels
# df_single %>%
#   st_drop_geometry() %>%
# group_by(damage_type_list) %>%
#   summarize(count=n(),
#             damage_type_label=toString(unlist(unique(damage_type_list)))) %>%
#   ungroup() %>%
#   select(damage_type_label, count) %>%
#   arrange(count)
# Mostly 
# Affected (1-9%), No Damage	182			
# Affected (1-9%), Destroyed (>50%)	216			
# Destroyed (>50%), No Damage	387	

# revise factor level for this map and others from now on -- using original dataframe for mapping that is ungrouped so we can see locations and distinctions
damaged_parcels_simple$damage_ordered <- factor(
      damaged_parcels_simple$damage,
      levels = c("Inaccessible","No Damage", "Affected (1-9%)", "Minor (10-25%)", "Major (26-50%)", "Destroyed (>50%)"), # Define the order of levels
      labels = c("Inaccessible","No Damage", "Affected","Minor","Major","Destroyed")
    )

pal <- colorFactor(palette = c("#6F6F6F","#909078","#177FEB", "#4886BD", "#CC983F","#FF9E0D"), domain = damaged_parcels_simple$damage_ordered)


# filter for these properties single resident properties
df_map <- damaged_parcels_simple %>%
  filter(apn %in% df_single$apn)

leaflet () %>%
addProviderTiles(providers$CartoDB.Positron) %>%
    setView(-118.104631, 34.185104, zoom = 12) %>%
  addPolygons(data=eaton_fire,
              fillColor="white",
              fillOpacity=0,
              color=meteorite,
              weight=2.5,
              opacity=1,
              popup=~"Eaton Fire Burn Area",
              group="Eaton Fire Perimeter") %>%
      addPolygons(data=lac_places_subset,
              fillColor="white",
              fillOpacity=0,
              color=peridot,
              weight=1.5,
              opacity=1,
              popup=~name,
              group="Cities") %>%
      addCircleMarkers(data=df_map,
        lng = ~longitude,
        lat = ~latitude,
        fillColor = ~pal(damage_ordered), # Color based on category
        radius = 2,
        stroke=TRUE,
        # fill=TRUE,
        fillOpacity = 1,
        color=gainsboro,
        weight=.5,
        opacity=.5,
        popup = ~paste0(damage,"</br>",
                        "Parcel #: ", apn, "</br>",
                        "Structure Category: ", structurec, "</br>",
                        "Address: ", full_address, "</br>",
                        "City: ", city),
        group="Single Residences with Multiple Damage Levels"
      ) %>%
      addLegend(
        position = "bottomright",
        pal = pal,
        values = df_map$damage_ordered,
        title = "Damage level"
      ) %>%
  addLayersControl(
          overlayGroups = c("Eaton Fire Perimeter", "Single Residences with Multiple Damage Levels","Cities"),
        options = layersControlOptions(collapsed = FALSE)
  )


```

I randomly selected 5 properties from the list to explore their damage combinations. Most even with only affected damage still have misfortune or calamity status approved for their tax bills. This is a status that property owners can apply for to receive a temporary reduction on their property taxes.

```{r}
# randomly select some parcels to test
sample <- sample_n(df_single,5)
# retrieved parcels below - this will produce a random list of 5 each time
# qa should focus on five people, not trying to reproduce the same 5

```

* Parcel Number: 5835017027
  Notes: A standing house with destroyed structures behind it and calamity status approved.
  https://portal.assessor.lacounty.gov/parceldetail/5835017027

* Parcel Number: 5840022039
  Notes: Property looks completely destroyed though it's only marked as Affected and No Damage. Calamity status is approved.
  https://portal.assessor.lacounty.gov/parceldetail/5840022039

* Parcel Number: 5857014006
  Notes: Unclear what structured on the property has been affected. Calamity status approved. Marked as Affected and No Damage.
  https://portal.assessor.lacounty.gov/parceldetail/5857014006

* Parcel Number: 5845031013
  Notes: Garage is destroyed. Unclear how house is damaged. Calamity status is approved.
  https://portal.assessor.lacounty.gov/parceldetail/5845031013

* Parcel Number: 5828008010
  Notes: Minor damage noted, but garage and house seem fully intact. House is a block from Lincoln Avenue. Calamity status is approved.
  https://portal.assessor.lacounty.gov/parceldetail/5828008010

## Multiple Residences with Multiple Damage Levels
This map plots the properties labeled as multiple residences that have multiple damage levels. A few that have no or affected damage are noteably outside of the fire perimeter.

```{r multiple-mixed}

# what's the target list for this map -- have to include other minor structure due to detached garages
target <- list("Multiple Residence", "Other Minor Structure")

# filter for parcels that have more than 2 damage levels and that are multiple residences
df_multiple <- unique_parcels %>%
  filter(damage_type_count>=2) %>%
  mutate(target_check = map_lgl(structure_category_list, ~any(!(. %in% target)))) %>%
  # filter for only parcels where category is in our target list--excludes any case where at least one category doesn't meet our list
  filter(target_check == FALSE)

# take out those only labels as other minor structure--they would have been mapped in previous one
df_multiple <- df_multiple %>%
  filter(!as.character(structure_category_list) == 'Other Minor Structure')


# # tabulate the combo damage levels
# df_multiple %>%
#   st_drop_geometry() %>%
# group_by(damage_type_list) %>%
#   summarize(count=n(),
#             damage_type_label=toString(unlist(unique(damage_type_list)))) %>%
#   ungroup() %>%
#   select(damage_type_label, count) %>%
#   arrange(count)
# Mostly 
# Destroyed (>50%), Major (26-50%)	3			
# Affected (1-9%), Destroyed (>50%), No Damage	3			
# Destroyed (>50%), No Damage	6			
# Affected (1-9%), No Damage	6	

# filter for these properties multiple resident properties
df_map <- damaged_parcels_simple %>%
  filter(apn %in% df_multiple$apn)


leaflet () %>%
addProviderTiles(providers$CartoDB.Positron) %>%
    setView(-118.104631, 34.185104, zoom = 12) %>%
  addPolygons(data=eaton_fire,
              fillColor="white",
              fillOpacity=0,
              color=meteorite,
              weight=2.5,
              opacity=1,
              popup=~"Eaton Fire Burn Area",
              group="Eaton Fire Perimeter") %>%
      addPolygons(data=lac_places_subset,
              fillColor="white",
              fillOpacity=0,
              color=peridot,
              weight=1.5,
              opacity=1,
              popup=~name,
              group="Cities") %>%
      addCircleMarkers(data=df_map,
        lng = ~longitude,
        lat = ~latitude,
        fillColor = ~pal(damage_ordered), # Color based on category
        radius = 2,
        stroke=TRUE,
        # fill=TRUE,
        fillOpacity = 1,
        color=gainsboro,
        weight=.5,
        opacity=.5,
        popup = ~paste0(damage,"</br>",
                        "Parcel #: ", apn, "</br>",
                        "Structure Category: ", structurec, "</br>",
                        "Address: ", full_address, "</br>",
                        "City: ", city),
        group="Multiple Residences with Multiple Damage Levels"
      ) %>%
      addLegend(
        position = "bottomright",
        pal = pal,
        values = df_map$damage_ordered,
        title = "Damage level"
      ) %>%
  addLayersControl(
          overlayGroups = c("Eaton Fire Perimeter", "Multiple Residences with Multiple Damage Levels","Cities"),
        options = layersControlOptions(collapsed = FALSE)
  )


```
I randomly selected 5 properties from the list to explore their damage combination. Most even with only affected damage still have misfortune and calamity status approved for their tax bills. Importantly, sometimes CalFire's database assigns the same parcel numbers to condominiums or multi-family properties that may have separate parcel numbers in L.A. County Assessor data. To join the two data sources, we will need to join points from CalFire's data to parcel polygons from the L.A. County Assessor, or join properties by a combination of address and parcel number.

```{r}
# randomly select some parcels to test
sample <- sample_n(df_multiple,5)
# retrieved parcels below
```

* Parcel Number: 5840015030
  Notes: Condo with a couple of destroyed buildings on it and calamity status approved. In this case, all 5 structures share the same parcel number.
  https://portal.assessor.lacounty.gov/parceldetail/5840015030

* Parcel Number: 5853008008
  Notes: Multi-family property with three units. Property is outside of the burn area. Calamity status is approved, but the units seem to be intact. They are noted as sustaining no or affected damage.
  https://portal.assessor.lacounty.gov/parceldetail/5853008008

* Parcel Number: 5845016006
  Notes: Multi-family property with three units. Calamity status is approved and the property is completely gone.
  https://portal.assessor.lacounty.gov/parceldetail/5845016006

* Parcel Number: 5751008029
  Notes: Condo complex with one building destroyed. Calamity status is approved, and property is delinquent on taxes. Complex is all on one parcel.
  https://portal.assessor.lacounty.gov/parceldetail/5751008029

* Parcel Number: 5844001010
  Notes: Multi-family property with one building destroyed and calamity status approved.
  https://portal.assessor.lacounty.gov/parceldetail/5844001010 


# Proposed Damage Categories
To focus on residential properties that sustained greater damage and make the data findings more easily interpretable (by including fewer categories), we can group damage levels as follows:

* No Damage: No Damage, Inaccessible. Properties that have no damage recorded on any structure.

* Some Damage: Affected (1-9%) or Minor Damage (10-25%) alone or in combination with No Damage or Inaccessible.

* Significant Damage or Destroyed: Major Damage (26-50%) or Destroyed (>50%) alone or in combination with another category.

## No Damage
These properties sustained no damage. It includes "inaccessible" but only 40 properties total were rated as inaccessible.

```{r}

# what's the target list 
target <- list("No Damage", "Inaccessible")

# filter for parcels that only have these damage levels
df_no_damage <- df_any_res_all %>%
  mutate(target_check = map_lgl(damage_type_list, ~any(!(. %in% target)))) %>%
  # filter for only parcels where no other damage type turns up
  filter(target_check == FALSE)

# filter for these properties to map
df_map <- damaged_parcels_simple %>%
  filter(apn %in% df_no_damage$apn)

# unique parcel count
unique_count <- length(unique(df_map$apn))

```

A total of `r unique_count` unique parcels fall into this category.


```{r}
leaflet () %>%
addProviderTiles(providers$CartoDB.Positron) %>%
    setView(-118.104631, 34.185104, zoom = 12) %>%
  addPolygons(data=eaton_fire,
              fillColor="white",
              fillOpacity=0,
              color=meteorite,
              weight=2.5,
              opacity=1,
              popup=~"Eaton Fire Burn Area",
              group="Eaton Fire Perimeter") %>%
      addPolygons(data=lac_places_subset,
              fillColor="white",
              fillOpacity=0,
              color=peridot,
              weight=1.5,
              opacity=1,
              popup=~name,
              group="Cities") %>%
      addCircleMarkers(data=df_map,
        lng = ~longitude,
        lat = ~latitude,
        fillColor = ~pal(damage_ordered), # Color based on category
        radius = 2,
        stroke=TRUE,
        # fill=TRUE,
        fillOpacity = 1,
        color=gainsboro,
        weight=.5,
        opacity=.5,
        popup = ~paste0(damage,"</br>",
                        "Parcel #: ", apn, "</br>",
                        "Structure Category: ", structurec, "</br>",
                        "Address: ", full_address, "</br>",
                        "City: ", city),
        group="Residences with No Damage"
      ) %>%
      addLegend(
        position = "bottomright",
        pal = pal,
        values = df_map$damage_ordered,
        title = "Damage level"
      ) %>%
  addLayersControl(
          overlayGroups = c("Eaton Fire Perimeter", "Residences with No Damage","Cities"),
        options = layersControlOptions(collapsed = FALSE)
  )

```


## Some Damage
These properties with at least some residential use would be considered as sustaining some damage. They sustained at least affected damage (1-9%) and at most minor damage (10-25%). We could combine this category with no damage to make a some to no damage category, or make one single damage category (combining with significant damage or destroyed).

```{r}

# what's the target list 
target <- list("No Damage", "Inaccessible", "Affected (1-9%)","Minor (10-25%)")

# filter for parcels that only have these damage levels
df_some_damage <- df_any_res_all %>%
  mutate(target_check = map_lgl(damage_type_list, ~any(!(. %in% target)))) %>%
  # filter for only parcels where no other damage type turns up
  filter(target_check == FALSE)

# filter for these properties to map and exclude the ones from the first list of little to no damage
df_map <- damaged_parcels_simple %>%
  filter(apn %in% df_some_damage$apn) %>%
  filter(!apn %in% df_no_damage$apn) 

# unique parcel count
unique_count <- length(unique(df_map$apn))

```

A total of `r unique_count` unique parcels fall into this category.

```{r}
leaflet () %>%
addProviderTiles(providers$CartoDB.Positron) %>%
    setView(-118.104631, 34.185104, zoom = 12) %>%
  addPolygons(data=eaton_fire,
              fillColor="white",
              fillOpacity=0,
              color=meteorite,
              weight=2.5,
              opacity=1,
              popup=~"Eaton Fire Burn Area",
              group="Eaton Fire Perimeter") %>%
      addPolygons(data=lac_places_subset,
              fillColor="white",
              fillOpacity=0,
              color=peridot,
              weight=1.5,
              opacity=1,
              popup=~name,
              group="Cities") %>%
      addCircleMarkers(data=df_map,
        lng = ~longitude,
        lat = ~latitude,
        fillColor = ~pal(damage_ordered), # Color based on category
        radius = 2,
        stroke=TRUE,
        # fill=TRUE,
        fillOpacity = 1,
        color=gainsboro,
        weight=.5,
        opacity=.5,
        popup = ~paste0(damage,"</br>",
                        "Parcel #: ", apn, "</br>",
                        "Structure Category: ", structurec, "</br>",
                        "Address: ", full_address, "</br>",
                        "City: ", city),
        group="Residences with Some Damage"
      ) %>%
      addLegend(
        position = "bottomright",
        pal = pal,
        values = df_map$damage_ordered,
        title = "Damage level"
      ) %>%
  addLayersControl(
          overlayGroups = c("Eaton Fire Perimeter", "Residences with Some Damage","Cities"),
        options = layersControlOptions(collapsed = FALSE)
  )

```

## Significant Damage or Destroyed
These properties with at least some residential use would be considered as sustaining significant damage or were destroyed. They sustained at least major damage (26-50%).

```{r}

# what's the target list 
target <- list("Destroyed (>50%)", "Major (26-50%)")

# filter for parcels that only have these damage levels
df_destroyed <- df_any_res_all %>%
  mutate(target_check = map_lgl(damage_type_list, ~any((. %in% target)))) %>%
  # filter for only parcels where at least one of these damage levels turns up
  filter(target_check == TRUE)

# filter for these properties to map 
df_map <- damaged_parcels_simple %>%
  filter(apn %in% df_destroyed$apn)

# unique parcel count
unique_count <- length(unique(df_map$apn))

```

A total of `r unique_count` unique parcels fall into this category.

```{r}

leaflet () %>%
addProviderTiles(providers$CartoDB.Positron) %>%
    setView(-118.104631, 34.185104, zoom = 12) %>%
  addPolygons(data=eaton_fire,
              fillColor="white",
              fillOpacity=0,
              color=meteorite,
              weight=2.5,
              opacity=1,
              popup=~"Eaton Fire Burn Area",
              group="Eaton Fire Perimeter") %>%
      addPolygons(data=lac_places_subset,
              fillColor="white",
              fillOpacity=0,
              color=peridot,
              weight=1.5,
              opacity=1,
              popup=~name,
              group="Cities") %>%
      addCircleMarkers(data=df_map,
        lng = ~longitude,
        lat = ~latitude,
        fillColor = ~pal(damage_ordered), # Color based on category
        radius = 2,
        stroke=TRUE,
        # fill=TRUE,
        fillOpacity = 1,
        color=gainsboro,
        weight=.5,
        opacity=.5,
        popup = ~paste0(damage,"</br>",
                        "Parcel #: ", apn, "</br>",
                        "Structure Category: ", structurec, "</br>",
                        "Address: ", full_address, "</br>",
                        "City: ", city),
        group="Residences with Some Damage"
      ) %>%
      addLegend(
        position = "bottomright",
        pal = pal,
        values = df_map$damage_ordered,
        title = "Damage level"
      ) %>%
  addLayersControl(
          overlayGroups = c("Eaton Fire Perimeter", "Residences with Some Damage","Cities"),
        options = layersControlOptions(collapsed = FALSE)
  )

```

# Eaton Fire Boundary Compared to Damaged Properties, Cities, and Tract Boundaries
This map displays residential or mixed use properties that sustained at least an affected (1-9%) damage level. The purpose of this map is to confirm boundaries for analysis, e.g., West Altadena to include the burn area West of Lake Avenue and East Altadena/Pasadena to include the burn area East of Lake Avenue and West of Michillinda.

Kinneloa Mesa did not sustain any damage so it would largely not be included in the analysis. Notably, some properties marked as destroyed fall just slightly outside of the burn area. The burn area might intersect with the parcel boundaries - data we are in progress of receiving. Only parcels in the burn area or sustaining at least some damage would be included in the detailed analysis. We can choose to include or not include properties that sustained at least a little damage but are outside of the official burn area.

```{r altadena-map}

# map damaged parcels and check boundaries we want to use against tracts and places

# for purpose of boundary mapping, let's just map properties that were damaged, regardless if they are duplicates and let's focus on residential properties
only_damaged_parcels <- damaged_parcels_simple %>%
  filter(!damage %in% c("No Damage", "Inaccessible")) %>%
  filter(structurec %in% c("Mixed Commercial/Residential","Multiple Residence","Single Residence","Other Minor Structure"))

# revise factor level
only_damaged_parcels$damage_ordered <- factor(
      only_damaged_parcels$damage,
      levels = c("Affected (1-9%)", "Minor (10-25%)", "Major (26-50%)", "Destroyed (>50%)"), # Define the order of levels
      labels = c("Affected","Minor","Major","Destroyed")
    )

# Create a color palette based on the damage level
# table(only_damaged_parcels$damage,useNA='always') # get values again to assign colors

pal <- colorFactor(palette = c("#177FEB", "#4886BD", "#CC983F","#FF9E0D"), domain = only_damaged_parcels$damage_ordered)
# table(only_damaged_parcels$damage_ordered,useNA='always')

leaflet () %>%
addProviderTiles(providers$CartoDB.Positron) %>%
    setView(-118.104631, 34.185104, zoom = 12) %>%
  addPolygons(data=eaton_fire,
              fillColor="white",
              fillOpacity=0,
              color=meteorite,
              weight=2.5,
              opacity=1,
              popup=~"Eaton Fire Burn Area",
              group="Eaton Fire Perimeter") %>%
    addPolygons(data=altadena_tracts,
              fillColor="white",
              fillOpacity=0,
              color=black,
              weight=1,
              opacity=1,
              popup=~name,
              group="Census Tracts") %>%
      addPolygons(data=lac_places_subset,
              fillColor="white",
              fillOpacity=0,
              color=peridot,
              weight=1.5,
              opacity=1,
              popup=~name,
              group="Cities") %>%
      addCircleMarkers(data=only_damaged_parcels,
        lng = ~longitude,
        lat = ~latitude,
        fillColor = ~pal(damage_ordered), # Color based on category
        radius = 2,
        stroke=TRUE,
        # fill=TRUE,
        fillOpacity = 1,
        color=gainsboro,
        weight=.5,
        opacity=.5,
        popup = ~paste0(damage,"</br>",
                        "Parcel #: ", apn, "</br>",
                        "Structure Category: ", structurec, "</br>",
                        "Address: ", full_address, "</br>",
                        "City: ", city),
                        # Optional: Add a popup
        group="Damaged Properties"
      ) %>%
      addLegend(
        position = "bottomright",
        pal = pal,
        values = only_damaged_parcels$damage_ordered,
        title = "Damage level"
      ) %>%
  addLayersControl(
          overlayGroups = c("Eaton Fire Perimeter", "Census Tracts","Damaged Properties","Cities"),
        options = layersControlOptions(collapsed = FALSE)
  )
      



```