---
title: "Commercial-Multiuse-Exploration"
output: html_document
---


```{r}
#load libraries
library(dplyr)
library(sf)
library(mapview)

#Add database connection and source script with functions for pushing to postgres
source("W:\\RDA Team\\R\\credentials_source.R")
con <- connect_to_db("altadena_recovery_rebuild")
options(scipen = 999)
```

```{r}
# Get the observed use codes in the Jan 2025 assessor data and April 2026
# Note: these have already been filtered for AINs in Altadena
get_use_code_query1 <- "SELECT 
    use_code, 
    COUNT(*) AS frequency
FROM 
    dashboard.assessor_data_universe_2025_01
GROUP BY 
    use_code
ORDER BY 
    frequency DESC;"

assessor_use_codes_jan_2025 <- dbGetQuery(con, get_use_code_query1) # 266
assessor_use_codes_jan_2025

get_use_code_query2 <- "SELECT 
    use_code, 
    COUNT(*) AS frequency
FROM 
    dashboard.assessor_data_universe_2026_04
GROUP BY 
    use_code
ORDER BY 
    frequency DESC;"

assessor_use_codes_apr_2026 <- dbGetQuery(con, get_use_code_query2) # 101
assessor_use_codes_apr_2026


# Data dictionary is here:W:\\Project\\RDA Team\\Altadena Recovery and Rebuild\\Data\\Assessor Data Extract\\FIELD DEF -- SBF.html
# Relevant codes summarized below

```

Based on the property use classification chart commercial use codes are 4 characters and generally start with a 1 or 2. Based on the data dictionary there are specific codes to note residential units on the property. In addition to this, the code can end in "C" to represent "condominiums." Here are the specific codes that site residential units:
- 111C
- 121*
- 1**C
- 171C
- 172*
- 184*
- 185*
- 191C
- 192C
- 193C
- 194C
- 2**C


Use codes related to residence but not included:
- Artist in residence: 102*
- Hotel under 50 rooms: 180*
- Hotel 50 rooms and over: 181*
- Motel under 50 rooms: 182*
- Motel 50 rooms and over: 183*

```{r}
# get commerical use codes of interest in jan 2025 and apr 2026
# use the codes to filter the data and parcel tables

### jan 2025
commercial_codes_jan_2025 <- assessor_use_codes_jan_2025 %>% 
  filter( 
    grepl("^1|^2", use_code)) %>% # 113
  mutate(
    commercial = TRUE,
    residential = case_when(
      grepl("C$", use_code) ~ TRUE,
      grepl("^121", use_code) ~ TRUE,
      grepl("^172", use_code) ~ TRUE,
      grepl("^184", use_code) ~ TRUE,
      grepl("^185", use_code) ~ TRUE,
      .default = FALSE))

# get all AINs 
get_jan_2025_data_query <- "SELECT * FROM dashboard.assessor_data_universe_2025_01;"
jan_2025_data <- dbGetQuery(con, get_jan_2025_data_query)

jan_2025_data_filtered <- commercial_codes_jan_2025 %>%
  left_join(jan_2025_data, by = "use_code") # 2629

# get shapes for these AINS
get_jan_2025_shapes_query <- "SELECT ain, matched_name, geom FROM dashboard.assessor_parcels_universe_2025_01;"
jan_2025_shapes <- st_read(con, query=get_jan_2025_shapes_query)

jan_2025_shapes_filtered <- jan_2025_shapes %>% 
  filter(ain %in% jan_2025_data_filtered$ain) %>% # 292
  filter(matched_name == "Altadena" | matched_name=="Pasadena; Altadena") %>% # 319
  left_join(jan_2025_data_filtered, by = c("ain"))

m_0125 <- jan_2025_shapes_filtered

# get list of ains to compare against apr 2026 vintage
jan_2025_ains <- jan_2025_data_filtered %>% filter(ain %in% jan_2025_shapes_filtered$ain)

### apr 2026
commercial_codes_apr_2026 <- assessor_use_codes_apr_2026 %>% 
  filter( 
    grepl("^1|^2", use_code)) %>% # 113
  mutate(
    commercial = TRUE,
    residential = case_when(
      grepl("C$", use_code) ~ TRUE,
      grepl("^121", use_code) ~ TRUE,
      grepl("^172", use_code) ~ TRUE,
      grepl("^184", use_code) ~ TRUE,
      grepl("^185", use_code) ~ TRUE,
      .default = FALSE
  ))

# get all AINs 
get_apr_2026_data_query <- "SELECT * FROM dashboard.assessor_data_universe_2026_04;"
apr_2026_data <- dbGetQuery(con, get_apr_2026_data_query)

apr_2026_data_filtered <- commercial_codes_apr_2026 %>%
  left_join(apr_2026_data, by = "use_code") # 2629

# get shapes for these AINS
get_apr_2026_shapes_query <- "SELECT ain, matched_name, geom FROM dashboard.assessor_parcels_universe_2026_04;"
apr_2026_shapes <- st_read(con, query=get_apr_2026_shapes_query)

apr_2026_shapes_filtered <- apr_2026_shapes %>% 
  filter(ain %in% apr_2026_data_filtered$ain) %>% # 47
  filter(matched_name == "Altadena" | matched_name=="Pasadena; Altadena") %>% # 315
  left_join(apr_2026_data_filtered, by = c("ain"))

m_0426 <- apr_2026_shapes_filtered

# get list of ains to compare against jan 2025 vintage
apr_2026_ains <- apr_2026_data_filtered %>% filter(ain %in% apr_2026_shapes_filtered$ain)

```

# January 2025 Commercial Properties
```{r}
mapview(m_0125, zcol="residential") 
```

# Apr 2026 Commercial Properties
```{r}
mapview(m_0426, zcol="residential") 
```

```{r}

# Note these are roughly the same - by count Jan 2025 has 48 AINs to Apr 2026 47

check1 <- setdiff(jan_2025_ains$ain, apr_2026_ains$ain)
check1 # "5828026005" "5845018001" "5845010020" "5828027024"

check2 <- setdiff(apr_2026_ains$ain, jan_2025_ains$ain)
check2 # none

jan_2025_data_filtered %>% filter(ain == "5845018001") # use_code 1210 (store and residential combo)

# LAC Assessor Portal currently considers this purely residential: https://portal.assessor.lacounty.gov/parceldetail/5845018001
# No history or parcel change (or any indication of why these was use code 1210 previously and is now 0400)

# Additionally in EPIC-LA nothing to explain change in use code: https://epicla.lacounty.gov/energov_prod/SelfService/#/search?m=1&fm=1&ps=10&pn=1&em=true&st=5845018001
# There is a "Fire Rapid Assessment Inspection" (Status Red) requested on 2/20/25
# We have not reviewed these types of inspections in relation to Eaton Fire Recovery (should explore)

```
