---
title: "Commercial-Multiuse-Exploration"
output: html_document
---

# Step 0: Setup
```{r, include=FALSE}
#load libraries
library(dplyr)
library(sf)
library(mapview)

#Add database connection and source script with functions for pushing to postgres
source("W:\\RDA Team\\R\\credentials_source.R")
con <- connect_to_db("altadena_recovery_rebuild")
options(scipen = 999)
```

# Step 1a: Review use codes in the January 2025 Assessor data against April 2026 use codes
```{r}
# Get the observed use codes in the Jan 2025 assessor data and April 2026
# Note: these have already been filtered for AINs in Altadena
get_use_code_query1 <- "SELECT 
    use_code, 
    COUNT(*) AS frequency
FROM 
    dashboard.assessor_data_universe_2025_01
GROUP BY 
    use_code
ORDER BY 
    frequency DESC;"

assessor_use_codes_jan_2025 <- dbGetQuery(con, get_use_code_query1) # 266
assessor_use_codes_jan_2025

get_use_code_query2 <- "SELECT 
    use_code, 
    COUNT(*) AS frequency
FROM 
    dashboard.assessor_data_universe_2026_04
GROUP BY 
    use_code
ORDER BY 
    frequency DESC;"

assessor_use_codes_apr_2026 <- dbGetQuery(con, get_use_code_query2) # 101
assessor_use_codes_apr_2026


# Data dictionary is here:W:\\Project\\RDA Team\\Altadena Recovery and Rebuild\\Data\\Assessor Data Extract\\FIELD DEF -- SBF.html
# Relevant codes summarized below

```

Based on the property use classification chart commercial use codes are 4 characters and generally start with a 1 or 2. Based on the data dictionary there are specific codes to note residential units on the property. In addition to this, the code can end in "C" to represent "condominiums." Here are the specific codes that site residential units:
- 111C
- 121*
- 1**C
- 171C
- 172*
- 184*
- 185*
- 191C
- 192C
- 193C
- 194C
- 2**C


Use codes related to residence but not included:
- Artist in residence: 102*
- Hotel under 50 rooms: 180*
- Hotel 50 rooms and over: 181*
- Motel under 50 rooms: 182*
- Motel 50 rooms and over: 183*

# Step 1b: Identify commercial and residential parcels in January 2025 and April 2026 data
```{r}
# get commerical use codes of interest in jan 2025 and apr 2026
# use the codes to filter the data and parcel tables

### jan 2025
commercial_codes_jan_2025 <- assessor_use_codes_jan_2025 %>% 
  filter( 
    grepl("^1|^2", use_code)) %>% # 113
  mutate(
    commercial = TRUE,
    residential = case_when(
      grepl("C$", use_code) ~ TRUE,
      grepl("^121", use_code) ~ TRUE,
      grepl("^172", use_code) ~ TRUE,
      grepl("^184", use_code) ~ TRUE,
      grepl("^185", use_code) ~ TRUE,
      .default = FALSE))

# get all AINs 
get_jan_2025_data_query <- "SELECT * FROM dashboard.assessor_data_universe_2025_01;"
jan_2025_data <- dbGetQuery(con, get_jan_2025_data_query)

jan_2025_data_filtered <- commercial_codes_jan_2025 %>%
  left_join(jan_2025_data, by = "use_code") # 2629

# get shapes for these AINS
get_jan_2025_shapes_query <- "SELECT ain, matched_name, geom FROM dashboard.assessor_parcels_universe_2025_01;"
jan_2025_shapes <- st_read(con, query=get_jan_2025_shapes_query)

jan_2025_shapes_filtered <- jan_2025_shapes %>% 
  filter(ain %in% jan_2025_data_filtered$ain) %>% # 292
  filter(matched_name == "Altadena" | matched_name=="Pasadena; Altadena") %>% # 319
  left_join(jan_2025_data_filtered, by = c("ain"))

m_0125 <- jan_2025_shapes_filtered

# get list of ains to compare against apr 2026 vintage
jan_2025_ains <- jan_2025_data_filtered %>% filter(ain %in% jan_2025_shapes_filtered$ain)

### apr 2026
commercial_codes_apr_2026 <- assessor_use_codes_apr_2026 %>% 
  filter( 
    grepl("^1|^2", use_code)) %>% # 113
  mutate(
    commercial = TRUE,
    residential = case_when(
      grepl("C$", use_code) ~ TRUE,
      grepl("^121", use_code) ~ TRUE,
      grepl("^172", use_code) ~ TRUE,
      grepl("^184", use_code) ~ TRUE,
      grepl("^185", use_code) ~ TRUE,
      .default = FALSE
  ))

# get all AINs 
get_apr_2026_data_query <- "SELECT * FROM dashboard.assessor_data_universe_2026_04;"
apr_2026_data <- dbGetQuery(con, get_apr_2026_data_query)

apr_2026_data_filtered <- commercial_codes_apr_2026 %>%
  left_join(apr_2026_data, by = "use_code") # 2629

# get shapes for these AINS
get_apr_2026_shapes_query <- "SELECT ain, matched_name, geom FROM dashboard.assessor_parcels_universe_2026_04;"
apr_2026_shapes <- st_read(con, query=get_apr_2026_shapes_query)

apr_2026_shapes_filtered <- apr_2026_shapes %>% 
  filter(ain %in% apr_2026_data_filtered$ain) %>% # 47
  filter(matched_name == "Altadena" | matched_name=="Pasadena; Altadena") %>% # 315
  left_join(apr_2026_data_filtered, by = c("ain"))

m_0426 <- apr_2026_shapes_filtered

# get list of ains to compare against jan 2025 vintage
apr_2026_ains <- apr_2026_data_filtered %>% filter(ain %in% apr_2026_shapes_filtered$ain)


table(jan_2025_ains$residential, useNA="ifany")
  
table(apr_2026_ains$residential, useNA="ifany")

```
In January 2025, there were 319 significantly damaged parcels with commercial use codes. Of those, 48 were also residential.

In April 2026, there were 315 significantly damage parcels with commercial use codes. Of those, 47 were also residential.


# Step 1C: View January 2025 and Apr 2026 Commercial Properties
```{r}
mapview(m_0125, zcol="residential") 

mapview(m_0426, zcol="residential") 
```

# Step 2a: Identify which properties are also zoned for and residential and commercial mixed-use purposes

In January 2025 there are two zoning codes that appear in comm/res mixed-use coded properties that don't appear in other use code types: 

- LCC2
- LCC3 R2-P

Common zoning codes in both lists:

- LCC2*
- LCC2YY
- LCC3*
- LCC3VV
- LCC3YY
- LCM1
- LCM1YY
- LCR175
- LCR2VV
- LCR3YY
- PSC- (should exclude, specific to Pasadena)

Repeating for April 2026 - again, two zoning codes in our filtered list of comm/res mixed use properties

- LCC2
- LCC3 R2-P

Common to both

- LCC2*
- LCC2YY
- LCC3*
- LCC3VV
- LCC3YY
- LCM1*
- LCM1YY
- LCR175
- LCR2VV
- LCR3YY
- PSC- (should exclude, specific to Pasadena)

Decision: Combine these lists excluding the code for Pasadena. Will keep LCM codes even though they suggest manufacturing is the primary use.

```{r}
# Get zoning codes to try and identify which commercial use codes are also zoned
# for residential (mixed-use) purposes.

# January 2025
zoning_codes_0125 <- jan_2025_data %>% filter(!(ain %in% jan_2025_ains$ain)) %>% select(zoning_code)
zoning_codes_0125_df <- as.data.frame(table(zoning_codes_0125$zoning_code))

zoning_codes_0125_filtered <- jan_2025_shapes_filtered %>% st_drop_geometry() %>% 
  filter(commercial==TRUE & residential == TRUE) %>% 
  select(zoning_code)

zoning_codes_0125_filtered_df <- as.data.frame(table(zoning_codes_0125_filtered$zoning_code))

# Review all observed zoning codes against those associated with residential and commercial mixed use properties
setdiff(as.character(zoning_codes_0125_df$Var1), as.character(zoning_codes_0125_filtered_df$Var1))
setdiff(as.character(zoning_codes_0125_filtered_df$Var1), as.character(zoning_codes_0125_df$Var1))
intersect(as.character(zoning_codes_0125_filtered_df$Var1), as.character(zoning_codes_0125_df$Var1))


# April 2026
zoning_codes_0426 <- apr_2026_data %>% filter(!(ain %in% apr_2026_ains$ain)) %>% select(zoning_code)
zoning_codes_0426_df <- as.data.frame(table(zoning_codes_0426$zoning_code))

zoning_codes_0426_filtered <- apr_2026_shapes_filtered %>% st_drop_geometry() %>% 
  filter(commercial==TRUE & residential == TRUE) %>% 
  select(zoning_code)

zoning_codes_0426_filtered_df <- as.data.frame(table(zoning_codes_0426_filtered$zoning_code))

# Review all observed zoning codes against those associated with residential and commercial mixed use properties
setdiff(as.character(zoning_codes_0426_df$Var1), as.character(zoning_codes_0426_filtered_df$Var1))
setdiff(as.character(zoning_codes_0426_filtered_df$Var1), as.character(zoning_codes_0426_df$Var1))
intersect(as.character(zoning_codes_0426_filtered_df$Var1), as.character(zoning_codes_0426_df$Var1))


# Same universe of zoning codes are available. Will review individually.
# Reference data dictionary
## First two characters are place (LC: LA County; PS: Pasadena)
## 3rd character: can be (A)gricultural, (C)ommercial, (M) Industrial, R(esidential)
## 4th character: alphanumeric and generally represents either the intensity or limit of a property's use;
## 5th-15th characters: represent data supplied and used by the Real Estate Division of the L.A. Office of Assessor 
## in the case of a dash, it is used to separate multiple zones or height districts.

# Questions:
# What do asterisks mean? (still not sure)

zoning_codes <- unique(c(
  setdiff(as.character(zoning_codes_0125_filtered_df$Var1), as.character(zoning_codes_0125_df$Var1)),
  intersect(as.character(zoning_codes_0125_filtered_df$Var1), as.character(zoning_codes_0125_df$Var1)),
  setdiff(as.character(zoning_codes_0426_filtered_df$Var1), as.character(zoning_codes_0426_df$Var1)),
  intersect(as.character(zoning_codes_0426_filtered_df$Var1), as.character(zoning_codes_0426_df$Var1))
  ))

```

Will only review with Apr 2026 data:
```{r}
apr_2026_zoned <- apr_2026_data %>% filter(zoning_code %in% zoning_codes)
zoning_freq <- as.data.frame(table(apr_2026_zoned$zoning_code, useNA="ifany"))
zoning_freq

apr_2026_lcr175 <- apr_2026_zoned %>% filter(grepl("^LCR175", zoning_code))
table(apr_2026_lcr175$use_code, useNA = "ifany")
# https://portal.assessor.lacounty.gov/parceldetail/5827013902
# 2700 is a parking lot - there are 4

# https://portal.assessor.lacounty.gov/parceldetail/5823020021
# Split residential, commercial - there is 1

zoning_codes_final <- Filter(function(x) !any(grepl("^LCR175", x)), zoning_codes)

apr_2026_zoned_final <- apr_2026_data %>% filter(zoning_code %in% zoning_codes_final) # 532

```


There are a tremendous number of LCR175	(n=7809 ) which seems primarily a residential zone. Reviewing the use codes these are overwhelmingly residential. Will exclude it from the list of zoning codes.

# Step 3: With the broad list of properties zoned for comm/res mixed use, vet data availability across sources



