---
title: "Commercial-Multiuse-Exploration"
output: html_document
---


```{r}
#load libraries
library(dplyr)
library(sf)
library(mapview)

#Add database connection and source script with functions for pushing to postgres
source("W:\\RDA Team\\R\\credentials_source.R")
con <- connect_to_db("altadena_recovery_rebuild")
options(scipen = 999)
```

```{r}
# Get the observed use codes in the Jan 2025 assessor data - 
# Note: these have already been filtered for AINs in Altadena

get_use_code_query <- "SELECT 
    use_code, 
    COUNT(*) AS frequency
FROM 
    dashboard.assessor_data_universe_2025_01
GROUP BY 
    use_code
ORDER BY 
    frequency DESC;"

assessor_use_codes_jan_2025 <- dbGetQuery(con, get_use_code_query) # 266

assessor_use_codes_jan_2025

# Data dictionary is here:W:\\Project\\RDA Team\\Altadena Recovery and Rebuild\\Data\\Assessor Data Extract\\FIELD DEF -- SBF.html
# Relevant codes summarized below

```

Based on the property use classification chart commercial use codes are 4 characters and generally start with a 1 or 2. Based on the data dictionary there are specific codes to note residential units on the property. In addition to this, the code can end in "C" to represent "condominiums." Here are the specific codes that site residential units:
- 111C
- 121*
- 1**C
- 171C
- 172*
- 184*
- 185*
- 191C
- 192C
- 193C
- 194C
- 2**C


Use codes related to residence but not included:
- Artist in residence: 102*
- Hotel under 50 rooms: 180*
- Hotel 50 rooms and over: 181*
- Motel under 50 rooms: 182*
- Motel 50 rooms and over: 183*

```{r}
commercial_codes <- assessor_use_codes_jan_2025 %>% 
  filter( 
    grepl("^1|^2", use_code)) %>% # 113
  mutate(residential = case_when(
    grepl("C$", use_code) ~ TRUE,
    grepl("^121", use_code) ~ TRUE,
    grepl("^172", use_code) ~ TRUE,
    grepl("^184", use_code) ~ TRUE,
    grepl("^185", use_code) ~ TRUE,
    .default = FALSE
  ))

commercial_residential_codes <- commercial_codes %>% filter(residential == TRUE)

# get all AINs 
get_jan_2025_data_query <- "SELECT * FROM dashboard.assessor_data_universe_2025_01;"
jan_2025_data <- dbGetQuery(con, get_jan_2025_data_query)

jan_2025_data_filtered <- jan_2025_data %>% filter(use_code %in% commercial_residential_codes$use_code) # 292

# get shapes for these AINS
get_jan_2025_shapes_query <- "SELECT ain, matched_name, geom FROM dashboard.assessor_parcels_universe_2025_01;"
jan_2025_shapes <- st_read(con, query=get_jan_2025_shapes_query)

jan_2025_shapes_filtered <- jan_2025_shapes %>% 
  filter(ain %in% jan_2025_data_filtered$ain) %>% # 292
  filter(matched_name == "Altadena" | matched_name=="Pasadena; Altadena") # 48
m <- jan_2025_shapes_filtered
mapview(m, zcol="matched_name")
```
