---
title: "Change_Summary_March_2026"
output: html_document
---

Objective: Explore rebuild status changes from December 2025 to March 2026 and 
identify any methodology refinements we should make. Focus primarily on the 
16 properties that:
1. Jumped to rebuild complete (essentially skipping an In Construction phase; n=2)
2. Regressed from rebuild complete to some earlier phase (n=6)
3. Newly designated as rebuilt and did not jump over construction (n=8)

Key considerations:
- Will review each property individually for permits and CalFire DINS data
- May also review RedFin/Zillow type sites if also sold after Eaton
- What methodology refinements can we make if any?
- What impact would the methodology refinements have on other properties (if any)


Other considerations:
- From ANFS integration: 1578 Morada Pl, Altadena, CA 91001 has sold and is for sale 
again for ~ $2M. We don't consider it rebuilt - what are we missing?
see: https://www.zillow.com/homedetails/1578-Morada-Pl-Altadena-CA-91001/20918644_zpid/ 
- Stepping back from these specific cases should make sure the "new" properties
that didn't skip construction also look good


```{r}
# Add database connection and source script with functions for pushing to postgres
source("W:\\RDA Team\\R\\credentials_source.R")
library(tidyr)

con <- connect_to_db("altadena_recovery_rebuild")
options(scipen = 999)

# get summary of changes
sql_query <- "SELECT curr.*, prev.dashboard_label as prev_dashboard_label,
CASE 
        WHEN curr.dashboard_label = prev.dashboard_label THEN 'unchanged'
        WHEN curr.dashboard_label = 'Repairs or Rebuild Complete' AND prev.dashboard_label != 'Repairs or Rebuild Complete' THEN 'new rebuild complete'
        WHEN curr.dashboard_label != 'Repairs or Rebuild Complete' AND prev.dashboard_label = 'Repairs or Rebuild Complete' THEN 'reverted - QA to figure out why'
        ELSE 'something else?'
    END AS change_summary
FROM dashboard.rel_parcel_rebuild_status_2026_03 curr
LEFT JOIN dashboard.rel_parcel_rebuild_status_2025_12 prev ON curr.ain = prev.ain
WHERE prev.dashboard_label = 'Repairs or Rebuild Complete' OR curr.dashboard_label = 'Repairs or Rebuild Complete';"

rebuild_changes <- dbGetQuery(con=con, sql_query) %>%
  mutate(qa_group = case_when(
    change_summary=='new rebuild complete'& prev_dashboard_label == "In Construction" ~ "new",
    change_summary=='new rebuild complete' & prev_dashboard_label != "In Construction" ~ "jumped",
    change_summary=='reverted - QA to figure out why' ~ "reverted",
    change_summary=='unchanged' ~ "unchanged",
    .default = "something else"
  ))

calfire_query <- "SELECT calfire.*, xwalk.ain_2025_12, xwalk.status as xwalk_status FROM data.rel_assessor_damage_level calfire
left join dashboard.crosswalk_assessor_2025_09_12 xwalk ON calfire.ain = xwalk.ain_2025_01;"
calfire <- dbGetQuery(con, calfire_query)

sales_query <- "SELECT ain, sold_after_eaton, last_sale_date, sale_two_date, sale_three_date FROM dashboard.rel_assessor_sales_2025_12;"
sales <- dbGetQuery(con, sales_query)

# Get permit details
permit_query <- "SELECT wf.ain, wf.permit_number, wf.status as wf_status, wf.status_date as wf_status_date, det.applied_date, det.issued_date, det.expire_date, det.finalized_date, wf.workflow_item, det.type as permit_type, 
det.status as permit_status, det.project_name, det.project_name_href, 
gen.description as permit_description
FROM dashboard.scraped_workflow_permit_data_2026_03 wf
left join dashboard.scraped_detailed_permit_data_2026_03 det on wf.ain = det.ain and wf.permit_number = det.permit_number
left join dashboard.scraped_general_permit_data_2026_03 gen on wf.ain = gen.ain and wf.permit_number = gen.permit_number;"

permits <- dbGetQuery(con, permit_query)

dbDisconnect(con)

# Create one df with minimal columns and attach CalFire DINS and Sales data
df <- rebuild_changes %>%
  select(ain, starts_with("b2_"), total_permits, starts_with("bucket_"), qa_group) %>%
  left_join(calfire, by=c("ain"="ain_2025_12"), suffix = c("", "_dmg")) %>%
  left_join(sales, by="ain", suffix = c("", "_sales"))


```

Joining the 33 properties to CalFire and sales data to get a fuller picture. 

Can see that only one has been sold. Confirmed against zillow that the home last sold in
March 2025 ($553k) and is on the market again after full rebuild keeping original
footprint/foundation ($1.649M). It's been on the market for 21 days (listed ~2/20/26). 
reference: https://www.zillow.com/homedetails/3010-Highview-Ave-Altadena-CA-91001/67088026_zpid/

```{r}
df %>% filter(sold_after_eaton==TRUE) %>% select(ain, sold_after_eaton, qa_group)

```

Sales data is less likely to inform the rest so focusing next on the "jumped" category

## Jumped: The 2 new rebuilds not previously "In Construction"

Below is a summary of what was reviewed:
- 5845014026 (prev. With Permit)
  - Assessor: https://portal.assessor.lacounty.gov/parceldetail/5845014026 (damage not discernible from aerial view)
  - Address: 958 MORADA PL, ALTADENA CA 91001-3114
  - Damage: 2 structures; 1 major, 1 none
  - Permits: 
    - Permit descriptions indicate the garage was likely the structure with major damage
    - Finalized dates suggest the garage was rebuilt/repaired on 1/14/26 with relevant electrical work finalized on 1/15/26
  - Conclusion: The jump checks out. At the time of December scrape there were no inspections on file. First one appears 1/07/26
- 5845015007 (prev. Without Permit)

Both have damage data stating multiple structures were on the property; one structure had no damage and the remaining were destroyed or had major damage

```{r}
jumped <- df %>% filter(qa_group=="jumped")
jumped %>%
  select(ain, structure_count, damage_type_list, ends_with("_damage_count"))
 
```

Review permits for these two parcels to see if we can learn more about which structures were damaged.

We can see 5845014026 


```{r}
jumped <- df %>% filter(qa_group=="jumped")
jumped %>%
  select(ain, structure_count, damage_type_list, ends_with("_damage_count")) %>%
  left_join(permits, by="ain") %>%
  View()
```

## The 8 new rebuilds that were prev. "In Construction" - new
- 5842003002 
- 5845005020 
- 5846006027 
- 5832014001 
- 5832017002 
- 5843019018
- 5833017001 - confirmed rebuilt via Zillow: https://www.zillow.com/homedetails/3010-Highview-Ave-Altadena-CA-91001/67088026_zpid/
- 5846022028

## The 6 AINs that reverted from rebuild complete - reverted
- 5844028003 (now "With Permit")
- 5846017025 (now "In Construction")
- 5846021004 (now "In Construction")
- 5828018005 (now "In Construction")
- 5829018039 (now "In Construction")
- 5829020023 (now "In Construction")


## There are 17 that remained rebuild completein Dec and March - unchanged 
